---
title: "ANOVA table"
output: html_document
editor_options:
  markdown:
    canonical: false
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, message = NA)
```

```{r LoadPackages, include=F}

# install the packages below if needed
pkgs <- c("knitr", "exams", "magrittr", "kableExtra", "tibble", "dplyr", "ggplot2", "car")
lapply(pkgs, library, character.only=T)

```

```{r SuppressScientificNotation}
options(scipen=999)
```

```{r DataGeneration, results = "hide"}
# DATA AND DATA-OBJECT GENERATION
Humb <- tibble(`Message Received`="Humblebrag", `Likeability of Sender` = rnorm(sample(c(30:40),1), 3.32*100, 1.23*100))
Brag <- tibble( `Message Received`="Brag", `Likeability of Sender` = rnorm(sample(c(30:40),1), 3.99*100, 1.28*100), )
Comp <- tibble( `Message Received`="Complaint", `Likeability of Sender` = rnorm(sample(c(30:40),1), 4.24*100, 0.88*100))

tib_X100 <- bind_rows(Humb, Brag, Comp)
tib <- tib_X100 %>% 
  mutate(`Likeability of Sender` = 
           case_when(`Likeability of Sender` > 600 ~ 7,
                     `Likeability of Sender` > 500 ~ 6,
                     `Likeability of Sender` > 400 ~ 5,
                     `Likeability of Sender` > 300 ~ 4,
                     `Likeability of Sender` > 200 ~ 3,
                     `Likeability of Sender` > 100 ~ 2,
                     TRUE ~ 1)
         )
rm(list=c("tib_X100", "Humb", "Brag", "Comp"))
```

```{r CreateDataSummary, results = "hide", message=F, warning=F}
# Create the data summary
GrpDesc <- tib %>% 
  group_by(`Message Received`) %>%
  summarize(n = length(`Likeability of Sender`), 
            M = mean(`Likeability of Sender`),
            SD = sd(`Likeability of Sender`),
            SE = SD/sqrt(n),
            `lower 95% CI` = M-1.96*SE,
            `upper 95% CI` = M+1.96*SE
            ) 
```

# Question

You replicate *Study 5* from Sezer, Gino, & Norton (2018; Study 5). The
original study had participants (*N* = 305) send each other text
messages that included either a *humblebrag* (i.e., "I am so exhausted
from getting elected to leadership positions"), or a *direct brag*
(i.e., "I get elected to leadership positions), or a *complaint*
(i.e.,"I am so exhausted"). The message receivers then rated how
*likeable*, *competent*, and *sincere* the senders seemed (1 = *not at
all* ... 7 = *very much*). Your replication is restricted to
`Likeability`. Descriptive statistics of each text-receiver group, along
with how they rated their respective text-message senders are in the
table and figure below.

------------------------------------------------------------------------

```{r DisplayDescriptivesTable, fig.cap='Descriptive statistics for *Likeability of Sender* as rated by the groups that received the different types of messages.'}
# Display the data summary in a table
GrpDesc %>% select(1:4) %>%
kable(digits=c(0,0,4,4), align = "lccc") %>%
  kable_styling(bootstrap_options = "striped")
```

------------------------------------------------------------------------

<br/>

**Calculations**

You run a oneway ANOVA on your replication data and get the following
*Sums of Squares* (*between* and *within*), but nothing else about the
ANOVA table that you need. Your task is to complete the table given what
you know about the study, the sum of squares in the table, and your
knowledge of the *F*-ratio. Each dash represents a statistic that you
will need to answer the questions further down. Notice that you do
***not*** need to calculate the *p*-value. At each stage of your
calculation, round to ***FOUR*** decimal places. But your final answer
should be rounded to ***THREE*** decimal places.

------------------------------------------------------------------------

```{r Model, results = "hide"}
# Build the model and format the ANOVA table
model <- lm(`Likeability of Sender` ~ `Message Received`, 
             data = tib,
             contrasts = list(`Message Received`="contr.sum"))
FTable <- car::Anova(model, type=3)

FTable$"Mean Squares" <- c("-", "-", "-")

options(knitr.kable.NA='')
colnames(FTable) <- c("Sum of Squares", "Degrees of Freedom", "F-ratio", "p-value", "Mean Squares")
rownames(FTable) <- c("(Intercept)", "Between", "Within")

FTable <- FTable %>% select(c(1,2,5,3,4))
```

```{r CreateSparseTable, results="hide"}
# blank out key statistics from the table
SparseFTable <- FTable
SparseFTable$`Degrees of Freedom` <- c(NA,"-","-")
SparseFTable$`F-ratio` <- c(NA,"-",NA)
SparseFTable$`p-value` <- c(NA,NA,NA)
```

```{r DisplaySparseTable}
# Display the blanked-out table
SparseFTable[-1,] %>% kable(
  digits=c(4,0,4,4,4), align = "ccccc", format="pandoc", caption = "ANOVA table"
  )
```

------------------------------------------------------------------------

```{r SeparateStatsObjects, results = "hide"}
# Create easily accessible names for different stats in order to make the markdown easier
SSb <- FTable$`Sum of Squares`[2]
SSb <- round(SSb,4)

SSw <- FTable$`Sum of Squares`[3]
SSw <- round(SSw,4)

DFb <- FTable$`Degrees of Freedom`[2]
DFw <- FTable$`Degrees of Freedom`[3]

MSb <- round(SSb/DFb,4)

MSw <- round(SSw/DFw,4)

Fratio <- round(MSb/MSw,3)

Pvalue <- FTable$`p-value`[2]
Pvalue <- round(Pvalue,4)

FCritVal <- qf(.05, DFb, DFw, lower.tail=FALSE)

G <- nlevels(as.factor(tib$`Message Received`))

N <- nrow(tib)
```

```{r QuestionVectorSize}
questions <- character(6)
solutions <- logical(6)
explanations <- character(6)
```

**TRUE/FALSE Questions**

Now that you have calculated these statistics, answer the questions
below. Check the box to indicate **TRUE**, and leave it blank to
indicate **FALSE**.

You will need to refer to [this online
table](https://patrickabolger.github.io/PSYC-301-Exam-Formulas/#f-table-alpha-05)
for the critical values of *F*.

You can also make reference to the [formulas and ANOVA tables for Unit
3](https://patrickabolger.github.io/PSYC-301-Exam-Formulas/#f-ratio-for-a-oneway-anova).

------------------------------------------------------------------------

<br/>

```{r QuestionGeneration}
# Question 1 (degrees of freedom between)

# Create false degrees of freedom between
fls_DFb <- sample(c(G-2,G,G+1,G+2),1)

# Generate Question 1
if(runif(1) > 0.5) {
  questions[1] <- paste0("The *degrees of freedom between* are ***less*** than $",fmt(fls_DFb, 0),"$.")
  solutions[1] <- DFb < fls_DFb
  explanations[1] <- paste0("The correct answer is ***",solutions[1],"***. The *degrees of freedom between* are the number of levels of the predictor variable minus 1 (*G*-1). Since there were $",G,"$ levels for the factor `Message Received`, the *degrees of freedom between* are $",fmt(DFb,0),"$. And $",fmt(DFb,0),"$ ",ifelse(solutions[1],"***is*** ", "is ***not*** "),"less than $",fmt(fls_DFb,0),"$.")
} else {
  questions[1] <- paste0("The *degrees of freedom between* are ***greater*** than $",fmt(fls_DFb, 0),"$.")
  solutions[1] <- DFb > fls_DFb
  explanations[1] <- paste0("The correct answer is ***",solutions[1],"***. The *degrees of freedom between* are the number of levels of the predictor variable minus 1 (*G*-1). Since there were $",G,"$ levels for the factor `Message Received`, the *degrees of freedom between* are $",fmt(DFb,0),"$. And $",fmt(DFb,0),"$ ",ifelse(solutions[1],"***is*** ", "is ***not*** "),"greater than $",fmt(fls_DFb,0),"$.")
  }

# Question 2  (degrees of freedom within)

# Create false degrees of freedom within
fls_DFw <- sample(c((N-(G-2)),(N-(G-1)),(N-(G+1)),(N-(G+2)),(N-(G+3)),(N-(G+4))),1)

# Generate Question 2
if(runif(1) > 0.5) {
  questions[2] <- paste0("The *degrees of freedom within* are ***less*** than $",fmt(fls_DFw,0),"$.")
  solutions[2] <- DFw < fls_DFw
  explanations[2] <- paste0("The correct answer is ***",solutions[2],"***. The *degrees of freedom within* are the total sample size (*N*) minus the number of levels in the predictor variable (*G*). So it is (*N*-*G*). Since the total sample size was $",fmt(N,0),"$, and there were $",fmt(G,0),"$ levels for the factor `Message Received`, the *degrees of freedom within* are $",fmt(DFw,0),"$. And $",fmt(DFw,0),"$ ",ifelse(solutions[2],"***is*** ", "is ***not*** "),"less than $",fmt(fls_DFw,0),"$.")
} else {
  questions[2] <- paste0("The *degrees of freedom within* are ***greater*** than $",fmt(fls_DFw,0),"$.")
  solutions[2] <- DFw > fls_DFw
  explanations[2] <- paste0("The correct answer is ***",solutions[2],"***. The *degrees of freedom within* are the total sample size (*N*) minus the number of levels in the predictor variable (*G*). So it is *N*-*G*. Since the total sample size was $",fmt(N,0),"$, and there were $",fmt(G,0),"$ levels for the factor `Message Received`, the *degrees of freedom within* are $",fmt(DFw,0),"$. And $",fmt(DFw,0),"$ ",ifelse(solutions[2],"***is*** ", "is ***not*** "),"greater than $",fmt(fls_DFw,0),"$.")
  }

# Question 3 (mean squares between)

# Create false Mean Squares between 
fls_MSb <- SSb/sample(c(DFb-1,DFb+1),1)

# Generate Question 3
if(runif(1) > 0.5) {
  questions[3] <- paste0("The *Mean Squares between* is ***less*** than $",fmt(fls_MSb,3),"$.")
  solutions[3] <- MSb < fls_MSb
  explanations[3] <- paste0("The correct answer is ***",solutions[3],"***. ","The *Mean Squares between* is the *Sum of Squares between* divided by the *degrees of freedom between*. Since the *Sum of Squares between* is $",fmt(SSb,3),"$, and the *degrees of freedom between* is $",fmt(DFb,0),"$, the *Mean Squares between* is $",fmt(MSb,3),"$. ","And $",fmt(MSb,3),"$ ",ifelse(solutions[3],"***is*** ", "is ***not*** "),"less than $",fmt(fls_MSb,3),"$.")
} else {
  questions[3] <- paste0("The *Mean Squares between* is ***greater*** than $",fmt(fls_MSb,3),"$.")
  solutions[3] <- MSb > fls_MSb
  explanations[3] <- paste0("The correct answer is ***",solutions[3],"***. ","The *Mean Squares between* is the *Sum of Squares between* divided by the *degrees of freedom between*. Since the *Sum of Squares between* is $",fmt(SSb,3),"$, and the *degrees of freedom between* is $",fmt(DFb,0),"$, the *Mean Squares between* is $",fmt(MSb,3),"$. ","And $",fmt(MSb,3),"$ ",ifelse(solutions[3],"***is*** ", "is ***not*** "),"greater than $",fmt(fls_MSb,3),"$.")
  }

# Question 4 (mean squares within)
fls_MSw <- SSw/sample(c(DFw-1,DFw+1),1)

# Generate Question 3
if(runif(1) > 0.5) {
  questions[4] <- paste0("The *Mean Squares within* is ***less*** than $",fmt(fls_MSw,3),"$.")
  solutions[4] <- MSw < fls_MSw
  explanations[4] <- paste0("The correct answer is ***",solutions[4],"***. ","The *Mean Squares within* is the *Sum of Squares within* divided by the *degrees of freedom within*. Since the *Sum of Squares within* is $",fmt(SSw,3),"$, and the *degrees of freedom within* is $",fmt(DFw,0),"$, the *Mean Squares within* is $",fmt(MSw,3),"$. ","And $",fmt(MSw,3),"$ ",ifelse(solutions[4],"***is*** ", "is ***not*** "),"less than $",fmt(fls_MSw,3),"$.")
} else {
  questions[4] <- paste0("The *Mean Squares within* is ***greater*** than $",fmt(fls_MSw,3),"$.")
  solutions[4] <- MSw > fls_MSw
  explanations[4] <- paste0("The correct answer is ***",solutions[4],"***. ","The *Mean Squares within* is the *Sum of Squares within* divided by the *degrees of freedom within*. Since the *Sum of Squares within* is $",fmt(SSw,3),"$, and the *degrees of freedom within* is $",fmt(DFw,0),"$, the *Mean Squares within* is $",fmt(MSw,3),"$. ","And $",fmt(MSw,3),"$ ",ifelse(solutions[4],"***is*** ", "is ***not*** "),"greater than $",fmt(fls_MSw,3),"$.")
  }
# Question 5 (The F-ratio)

fls_Fratio <- sample(c(Fratio-.01*Fratio, Fratio+.01*Fratio),1) 

if(runif(1) > 0.5) {
  questions[5] <- paste0("The *F*-ratio for this analysis is ***less*** than $",fmt(fls_Fratio,3),"$.")
  solutions[5] <- Fratio < fls_Fratio
  explanations[5] <- paste0("The correct answer is ***",solutions[5],"***. The *Means Squares between* for your replication experiment is $", fmt(MSb,3),"$, and the *Mean Squares within* is $", fmt(MSw,3), "$. Divide the former by the latter, and you get the *F*-ratio. In this case, it is $",fmt(Fratio,3),"$. and therefore, it ",ifelse(solutions[5],"***is*** ", "is ***not*** "),"less than $",fmt(fls_Fratio,3),"$.")
} else {
  questions[5] <- paste0("The *F*-ratio for this analysis is ***greater*** than $",fmt(fls_Fratio,3),"$.")
  solutions[5] <- Fratio > fls_Fratio
  explanations[5] <- paste0("The correct answer is ***",solutions[5],"***. The *Mean Squares between* for your replication experiment is $", fmt(MSb,3), "$, and the *Mean Squares within* is $", fmt(MSw,3), "$. Divide the former by the latter, and you get the *F*-ratio. In this case, it is $",fmt(Fratio,3),"$ and therefore, it ",ifelse(solutions[5],"***is*** ", "is ***not*** "),"greater than $",fmt(fls_Fratio,3),"$.")
  }

# Question 6 (reject or fail to reject null hypothesis)
# Get critical value

if(runif(1) > 0.5) {
  questions[6] <- "Given the *degrees of freedom between*, the *degrees of freedom within*, the *F*-ratio, and an *alpha* level of .05, your decision should be to ***reject*** the null hypothesis."
  solutions[6] <- Fratio >= FCritVal
  explanations[6] <- paste0("The correct answer is ***",solutions[6],"***. Your *F*-ratio of $",fmt(Fratio,3),"$ (with $",fmt(DFb,0),"$ *degrees of freedom between*, and $",fmt(DFw,0),"$ *degrees of freedom within*) is ",ifelse(solutions[6], "***greater*** than or equal to ", "***less*** than "),"the critical value of $",fmt(FCritVal,3),"$. So the appropriate decision is to ",ifelse(solutions[6], "***reject*** the null hypothesis.", "***fail to reject*** the null hypothesis."))
} else {
  questions[6] <- "Given the *degrees of freedom between*, the *degrees of freedom within*, the *F*-ratio, and an *alpha* level of .05, your decision should be to **fail to** reject the null hypothesis."
  solutions[6] <- Fratio < FCritVal
  explanations[6] <- paste0("The correct answer is ***",solutions[6],"***. Your *F*-ratio of $",fmt(Fratio,3),"$ (with $",fmt(DFb,0),"$ *degrees of freedom between*, and $",fmt(DFw,0),"$ *degrees of freedom within*) is ",ifelse(solutions[6], "***less*** than ","***greater*** than or equal to "),"the critical value of $",fmt(FCritVal,3),"$. So the appropriate decision is to ",ifelse(solutions[6], "***fail to reject*** the null hypothesis.", "***reject*** the null hypothesis."))
}
```

```{r ifAllFalseMakeOneTRUE}

i <- 0
if(all(solutions==FALSE)) {
  i <- sample.int(6,1)
  }

if(i==1) {
  solutions[1] <- TRUE
  if(DFb < fls_DFb) {
  questions[1] <- paste0("The *degrees of freedom between* are ***less*** than $",fmt(fls_DFb,0),"$.")
  explanations[1] <- paste0("The correct answer is ***TRUE***. The *degrees of freedom between* are the number of levels of the predictor variable minus 1 (*G*-1). Since there were $",G,"$ levels for the factor `Message Received`, the *degrees of freedom between* are $",fmt(DFb,0),"$, and hence, ***are*** less than $",fmt(fls_DFb,0),"$.")
  } else {
    questions[1] <- paste0("The *degrees of freedom between* are ***greater*** than $",fmt(fls_DFb,0),"$.")
    explanations[1] <- paste0("The correct answer is ***TRUE***. The *degrees of freedom between* are the number of levels of the predictor variable minus 1 (*G*-1). Since there were $",G,"$ levels for the factor `Message Received`, the *degrees of freedom between* are $",fmt(DFb,0),"$, and hence, ***are*** greater than $",fmt(fls_DFb,0),"$.")
  }
} else if(i==2) {
  solutions[2] <- TRUE
  if(DFw < fls_DFw) {
  questions[2] <- paste0("The *degrees of freedom within* are ***less*** than $",fmt(fls_DFw,0),"$.")
  explanations[2] <- paste0("The correct answer is ***TRUE***. The *degrees of freedom within* are the total sample size (*N*) minus the number of levels in the predictor variable (*G*). So it is (*N*-*G*). Since the total sample size was $",fmt(N,0),"$, and there were $",fmt(G,0),"$ levels for the factor `Message Received`, the *degrees of freedom within* are $",fmt(DFw,0),"$, and hence, ***are*** less than $",fmt(fls_DFw,0),"$.")
  } else {
    questions[2] <- paste0("The *degrees of freedom within* are ***greater*** than $",fmt(fls_DFw,0),"$.")
    explanations[2] <- paste0("The correct answer is ***TRUE***. The *degrees of freedom within* are the total sample size (*N*) minus the number of levels in the predictor variable (*G*). So it is (*N*-*G*). Since the total sample size was $",fmt(N,0),"$, and there were $",fmt(G,0),"$ levels for the factor `Message Received`, the *degrees of freedom within* are $",fmt(DFw,0),"$, and hence, ***are*** greater than $",fmt(fls_DFw,0),"$.")
  }
} else if (i==3) {
  solutions[3] <- TRUE
  if (MSb < fls_MSb) {
  questions[3] <- paste0("The *Mean Squares between* is ***less*** than $",fmt(fls_MSb,3),"$.")
  explanations[3] <- paste0("The correct answer is ***TRUE***. The *Mean Squares between* is the *Sum of Squares between* divided by the *degrees of freedom between*. Since the *Sum of Squares between* is $",fmt(SSb,3),"$, and the *degrees of freedom between* is $",fmt(DFb,0),"$, *the Mean Squares between* is $",fmt(MSb,3),"$. And $",fmt(MSb,3),"$ ***is*** less than $",fmt(fls_MSb,3),"$.")
} else {
  questions[3] <- paste0("The *Mean Squares between* is ***greater*** than $",fmt(fls_MSb,3),"$.")
  explanations[3] <- paste0("The correct answer is ***TRUE***. The *Mean Squares between* is the *Sum of Squares between* divided by the *degrees of freedom between*. Since the *Sum of Squares between* is $",fmt(SSb,3),"$, and the *degrees of freedom between* is $",fmt(DFb,0),"$, the *Mean Squares between* is $",fmt(MSb,3),"$. And $",fmt(MSb,3),"$ ***is*** greater than $",fmt(fls_MSb,3),"$.")
  }
} else if (i==4) {
  solutions[4] <- TRUE
  if (MSw < fls_MSw) {
  questions[4] <- paste0("The *Mean Squares within* is ***less*** than $",fmt(fls_MSw,3),"$.")
  explanations[4] <- paste0("The correct answer is ***TRUE***. The *Mean Squares within* is the *Sum of Squares within* divided by the *degrees of freedom within*. Since the *Sum of Squares within* is $",fmt(SSw,3),"$, and the *degrees of freedom within* is $",fmt(DFw,0),"$, the *Mean Squares within* is $",fmt(MSw,3),"$. And $",fmt(MSw,3),"$ ***is*** less than $",fmt(fls_MSw,3),"$.")
  } else {
  questions[4] <- paste0("The *Mean Squares within* is ***greater*** than $",fmt(fls_MSw,3),"$.")
  explanations[4] <- paste0("The correct answer is ***TRUE***. The *Mean Squares within* is the *Sum of Squares within* divided by the *degrees of freedom within*. Since the *Sum of Squares within* is $",fmt(SSw,3),"$, and the *degrees of freedom within* is $",fmt(DFw,0),"$, the *Mean Squares within* is $",fmt(MSw,3),"$. And $",fmt(MSw,3),"$ ***is*** greater than $",fmt(fls_MSw,3),"$.")
  }
} else if (i==5) {
  solutions[5] <- TRUE
  if (Fratio >= FCritVal) {
  questions[5] <- paste0("The *F*-ratio for this analysis is ***less*** than $",fmt(fls_Fratio,3),"$.")
  explanations[5] <- paste0("The correct answer is ***TRUE***. The *Means Squares between* for your replication experiment is $", fmt(MSb,3),"$, and the *Mean Squares within* is $", fmt(MSw,3), "$. Divide the former by the latter, and you get the *F*-ratio. In this case, it is $",fmt(Fratio,3),"$. and therefore, it ***is*** less than $",fmt(fls_Fratio,3),"$.")
  } else {
  questions[5] <- paste0("The *F*-ratio for this analysis is ***greater*** than $",fmt(fls_Fratio,3),"$.")
  explanations[5] <- paste0("The correct answer is ***TRUE***. The *Means Squares between* for your replication experiment is $", fmt(MSb,3),"$, and the *Mean Squares within* is $", fmt(MSw,3), "$. Divide the former by the latter, and you get the *F*-ratio. In this case, it is $",fmt(Fratio,3),"$. and therefore, it ***is*** greater than $",fmt(fls_Fratio,3),"$.")    
  }
} else if (i==6) {
  solutions[6] <- TRUE
  if (Fratio >= FCritVal) {
  questions[6] <- paste0("Given the *degrees of freedom between*, *the degrees of freedom within*, the *F*-ratio, and an *alpha* level of .05, your decision should be to ", ifelse(solutions[6], "***reject***", "***fail to reject***"), "$ the null hypothesis.")
  explanations[6] <- paste0("The correct answer is ***TRUE***. Your *F*-ratio of $",fmt(Fratio,3),"$ (with $",fmt(DFb,0),"$ *degrees of freedom between*, and $",fmt(DFw,0),"$ *degrees of freedom within*) is $", ifelse(solutions[6], "***greater than or equal to***", "***less*** than"), "$ the critical value of $",fmt(FCritVal,3),"$. So the appropriate decision is to $", ifelse(solutions[6], "***reject*** the null hypothesis", "***fail to reject*** the null hypothesis"),".")
} else {
  
  }
}

rm(i)

```

```{r ifAllTrueMakeOneFalse}

j <- 0
if(all(solutions==TRUE)) {
  j <- sample.int(6,1)
  }

if(j==1) {
  solutions[1] <- FALSE
  if(DFb > fls_DFb) {
  questions[1] <- paste0("The *degrees of freedom between* are ***less*** than $",fmt(fls_DFb,0),"$.")
  explanations[1] <- paste0("The correct answer is ***FALSE***. The *degrees of freedom between* are the number of levels of the predictor variable minus 1 (*G*-1). Since there were $",G,"$ levels for the factor `Message Received`, the *degrees of freedom between* are $",fmt(DFb,0),"$, and hence, are ***not*** less than $",fmt(fls_DFb,0),"$.")
  } else {
    questions[1] <- paste0("The *degrees of freedom between* are ***greater*** than $",fmt(fls_DFb,0),"$.")
    explanations[1] <- paste0("The correct answer is ***FALSE***. The *degrees of freedom between* are the number of levels of the predictor variable minus 1 (*G*-1). Since there were $",G,"$ levels for the factor `Message Received`, the *degrees of freedom between* are $",fmt(DFb,0),"$, and hence, are ***not*** greater than $",fmt(fls_DFb,0),"$.")
  }
} else if(j==2) {
  solutions[2] <- FALSE
  if(DFw > fls_DFw) {
  questions[2] <- paste0("The *degrees of freedom within* are ***less*** than $",fmt(fls_DFw,0),"$.")
  explanations[2] <- paste0("The correct answer is ***FALSE***. The *degrees of freedom within* are the total sample size (*N*) minus the number of levels in the predictor variable (*G*). So it is (*N*-*G*). Since the total sample size was $",fmt(N,0),"$, and there were $",fmt(G,0),"$ levels for the factor `Message Received`, the *degrees of freedom within* are $",fmt(DFw,0),"$, and hence, are ***not**** less than $",fmt(fls_DFw,0),"$.")
  } else {
    questions[2] <- paste0("The *degrees of freedom within* are ***greater*** than $",fmt(fls_DFw,0),"$.")
    explanations[2] <- paste0("The correct answer is ***FALSE***. The *degrees of freedom within* are the total sample size (*N*) minus the number of levels in the predictor variable (*G*). So it is (*N*-*G*). Since the total sample size was $",fmt(N,0),"$, and there were $",fmt(G,0),"$ levels for the factor `Message Received`, the *degrees of freedom within* are $",fmt(DFw,0),"$, and hence, are ***not*** greater than $",fmt(fls_DFw,0),"$.")
  }
} else if (j==3) {
  solutions[3] <- FALSE
  if (MSb > fls_MSb) {
  questions[3] <- paste0("The *Mean Squares between* is ***less*** than $",fmt(fls_MSb,3),"$.")
  explanations[3] <- paste0("The correct answer is ***FALSE***. The *Mean Squares between* is the *Sum of Squares between* divided by the *degrees of freedom between*. Since the *Sum of Squares between* is $",fmt(SSb,3),"$, and the *degrees of freedom between* is $",fmt(DFb,0),"$, *the Mean Squares between* is $",fmt(MSb,3),"$. And $",fmt(MSb,3),"$ is ***not*** less than $",fmt(fls_MSb,3),"$.")
} else {
  questions[3] <- paste0("The *Mean Squares between* is ***greater*** than $",fmt(fls_MSb,3),"$.")
  explanations[3] <- paste0("The correct answer is ***FALSE***. The *Mean Squares between* is the *Sum of Squares between* divided by the *degrees of freedom between*. Since the *Sum of Squares between* is $",fmt(SSb,3),"$, and the *degrees of freedom between* is $",fmt(DFb,0),"$, the *Mean Squares between* is $",fmt(MSb,3),"$. And $",fmt(MSb,3),"$ is ***not*** greater than $",fmt(fls_MSb,3),"$.")
  }
} else if (j==4) {
  solutions[4] <- FALSE
  if (MSw > fls_MSw) {
  questions[4] <- paste0("The *Mean Squares within* is ***less*** than $",fmt(fls_MSw,3),"$.")
  explanations[4] <- paste0("The correct answer is ***FALSE***. The *Mean Squares within* is the *Sum of Squares within* divided by the *degrees of freedom within*. Since the *Sum of Squares within* is $",fmt(SSw,3),"$, and the *degrees of freedom within* is $",fmt(DFw,0),"$, the *Mean Squares within* is $",fmt(MSw,3),"$. And $",fmt(MSw,3),"$ is ***not*** less than $",fmt(fls_MSw,3),"$.")
  } else {
  questions[4] <- paste0("The *Mean Squares within* is ***greater*** than $",fmt(fls_MSw,3),"$.")
  explanations[4] <- paste0("The correct answer is ***FALSE***. The *Mean Squares within* is the *Sum of Squares within* divided by the *degrees of freedom within*. Since the *Sum of Squares within* is $",fmt(SSw,3),"$, and the *degrees of freedom within* is $",fmt(DFw,0),"$, the *Mean Squares within* is $",fmt(MSw,3),"$. And $",fmt(MSw,3),"$ is ***not*** greater than $",fmt(fls_MSw,3),"$.")
  }
} else if (j==5) {
  solutions[5] <- FALSE
  if (Fratio > FCritVal) {
  questions[5] <- paste0("The *F*-ratio for this analysis is ***less*** than $",fmt(fls_Fratio,3),"$.")
  explanations[5] <- paste0("The correct answer is ***FALSE***. The *Mean Squares between* for your replication experiment is $", fmt(MSb,3),"$, and the *Mean Squares within* is $", fmt(MSw,3), "$. Divide the former by the latter, and you get the *F*-ratio. In this case, it is $",fmt(Fratio,3),"$. and therefore, it is ***not*** less than or equal to $",fmt(fls_Fratio,3),"$.")
  } else {
  questions[5] <- paste0("The *F*-ratio for this analysis is ***greater than or equal to*** $",fmt(fls_Fratio,3),"$.")
  explanations[5] <- paste0("The correct answer is ***FALSE***. The *Mean Squares between* for your replication experiment is $", fmt(MSb,3),"$, and the *Mean Squares within* is $", fmt(MSw,3), "$. Divide the former by the latter, and you get the *F*-ratio. In this case, it is $",fmt(Fratio,3),"$. and therefore, it is ***not*** greater than or equal to $",fmt(fls_Fratio,3),"$.")    
  }
} else if (j==6) {
  solutions[6] <- FALSE
  if (Fratio < FCritVal) {
  questions[6] <- paste0("Given the *degrees of freedom between*, the *degrees of freedom within*, the *F*-ratio, and an *alpha* level of .05, your decision should be to ", ifelse(solutions[6], "**reject**", "**fail to reject**"), "$ the null hypothesis.")
  explanations[6] <- paste0("The correct answer is ***FALSE***. Your *F*-ratio of $",fmt(Fratio,3),"$ (with $",fmt(DFb,0),"$ *degrees of freedom between*, and $",fmt(DFw,0),"$ *degrees of freedom within*) is $", ifelse(solutions[6], "***less than**", "***greater than or equal to***"), "$ the critical value of $",fmt(FCritVal,3),"$. So the appropriate decision is to $", ifelse(solutions[6], "***fail to reject*** the null hypothesis", "***reject*** the null hypothesis"),".")
} else {
  
  }
}

rm(j)

```

```{r questionlist, results = "asis"}
answerlist(questions, markup = "markdown")
```

# Solution

The full formula for the *F*-ratio is as follows:\
<br/>

$$F=\frac{Sum\ of\ Squares_{(between)}\bigg/degrees\ of\ freedom_{(between)}}{Sum\ of\ Squares_{(within)}\bigg/degrees\ of\ freedom_{(within)}}=\frac{MS_{between}}{MS_{within}}$$

------------------------------------------------------------------------

Or, in shorthand

$$F=\frac{SS_{between}\bigg/df_{between}}{SS_{within}\bigg/df_{within}}=\frac{MS_{between}}{MS_{within}}$$

------------------------------------------------------------------------

Or even shorter:

$$F=\frac{SS_b\bigg/df_b}{SS_w\bigg/df_w}$$

------------------------------------------------------------------------

<br/>

The following was true about this problem:

1.  You were given the *Sum of Squares* (both *between* and *within*) in
    the ANOVA table\
2.  You can calculate the *degrees of freedom between* from the
    knowledge of the study design alone\
3.  You can calculate the *degrees of freedom within* from the number of
    participants and the study design

This means that you were able fill out all the missing parts of the
table.

------------------------------------------------------------------------

<br/> The formula for the *degrees of freedom between* is $G-1$, or the
number of levels in the predictor variable minus 1. <br/>

Thus: <br/>

$$G-1=`r G`-1=`r G-1`$$

------------------------------------------------------------------------

<br/>

The formula for the *degrees of freedom within* is $N-G$, or the number
of participants minus the number of levels in the predictor variable.
<br/>

Thus: <br/>

$$N-G=`r N`-`r G`=`r N-G`$$

------------------------------------------------------------------------

<br/>

The formula for the *mean squares between* is $\frac{SS_b}{df_b}$, or
the *sum of squares between* divided by the *degrees of freedom
between*. <br/>

Thus: <br/>

$$MS_b=\frac{SS_b}{df_b}=\frac{`r SSb`}{`r DFb`}=`r round(SSb/DFb,3)`$$

------------------------------------------------------------------------

<br/>

The formula for the *mean squares within* is $\frac{SS_w}{df_w}$, or the
*sum of squares within* divided by the *degrees of freedom within*.
<br/>

Thus: <br/>

$$MS_w=\frac{SS_w}{df_w}=\frac{`r SSw`}{`r DFw`}=`r round(SSw/DFw,3)`$$

------------------------------------------------------------------------

<br/>

The formula for the *F*-ratio is $\frac{MS_b}{MS_w}$, or the *mean
squares between* divided by the *mean squares within*. <br/>

Thus: <br/>

$$F=\frac{MS_b}{MS_w}=\frac{`r MSb`}{`r MSw`}=`r round(MSb/MSw,3)`$$

------------------------------------------------------------------------

To get the critical value for an *F*-test with $\alpha=.05$, you go to
[this online
table](https://patrickabolger.github.io/PSYC-301-Exam-Formulas/#f-table-alpha-05)
and make the following choices:

-   Numerator degrees of freedom ($d_1$)
    -   `r DFb`
-   Denominator degrees of freedom ($d_2$)
    -   `r DFw`\
-   Then look at the intersection of those values for the critical value

------------------------------------------------------------------------

If you are interested, an error-bar chart of the data is below:

```{r, DisplayDescriptivesFigure}
# Represent the data in a error-bar chart
GrpDesc %>% ggplot(aes(x = `Message Received`)) + 
  geom_errorbar(aes(ymin=`lower 95% CI`, ymax=`upper 95% CI`), width = 0.4, color="#500000") +
  geom_point(aes(y=M), stat = "identity", color="#500000") +
  ylim(1,7) +
  labs(y="Likeability of Sender")
```

------------------------------------------------------------------------

***Reference***

Sezer, O, Gino, F., and Norton, M. I. (2018). Humblebragging: A
distinct - and ineffective - self-presentation strategy. *Journal of
Personality and Social Psychology*, *114*(1), 52-74.

```{r solutionlist, results = "asis"}
answerlist(explanations, markup = "markdown")
```

# Meta-information

extype: mchoice exsolution: `r mchoice2string(solutions)` exname: FTable
