---
title: "chisquare"
output: html_document
editor_options:
  markdown:
    canonical: false
    wrap: 72
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, message = NA)
```

```{r LoadPackages, include=F}

# install the packages below if needed
pkgs <- c("tibble", "dplyr", "forcats", "magrittr", "exams", "kableExtra", "ggplot2")
lapply(pkgs, library, character.only=T)

```

```{r SuppressScientificNotation}
options(scipen=999)
```

```{r data generation, echo = FALSE, message=F, warning=F, results = "hide"}
d01 <- c(
  rep("Freshman", sample(35:45,1)), 
  rep("Sophomore", sample(25:40,1)), 
  rep("Junior", sample(15:35,1)),
  rep("Senior", sample(15:35,1))
  )

d01 <- as_tibble(d01)
N <- nrow(d01)
d01 <- d01 %>% rename(`Class Level` = value)
d02 <- d01 %>% rename(`ClassLevel` = `Class Level`)

s01 <- d01 %>%
  group_by(`Class Level`) %>%
  summarize(`Count` = n())

s02 <- d02 %>%
  group_by(`ClassLevel`) %>%
  summarize(`Count` = n())

test <- chisq.test(table(d01$`Class Level`), p = c(1/4, 1/4, 1/4, 1/4))

# for plotting a pie chart
s03 <- s01 %>%
  group_by(`Class Level`) %>%
  ungroup() %>%
  mutate(percentage = round(`Count`/sum(`Count`),2)*100,
         lab.pos = cumsum(percentage)-.5*percentage)

```

# Question

You're the Head Director of a major student organization at Texas A&M. Out of curiosity, you wonder whether you have a balance of students across class levels. There are `r N` students in your organization. So you count the number of freshmen, sophomores, juniors, and seniors. Do you have an even balance?

The counts of students in each class level are listed below.

|     Freshmen     |    Sophomore     |      Junior      |      Senior      |
|:----------------:|:----------------:|:----------------:|:----------------:|
| `r s01$Count[1]` | `r s01$Count[4]` | `r s01$Count[2]` | `r s01$Count[3]` |

------------------------------------------------------------------------

**Calculations**

Calculate a chi-square *goodness-of-fit* test for these counts. At each stage of your calculations, round to ***FOUR*** decimal places. But your final answer should be rounded to ***THREE*** decimal places.

------------------------------------------------------------------------

**TRUE/FALSE Questions**

Now that you have calculated these statistics, answer the questions below. Check the box to indicate **TRUE**, and leave it blank to indicate **FALSE**.

You can use the following webpage for the critical values:

[Critical values for chi-square](https://patrickabolger.github.io/PSYC-301-Exam-Formulas/#chi2-table)

------------------------------------------------------------------------

<br/>

```{r QuestionVectorSize, echo=F, results = "hide"}
questions <- character(4)
solutions <- logical(4)
explanations <- character(4)
```

```{r QuestionGeneration, echo=F, results = "hide"}
# Create a false Expected Value, X2, and p value above or below the true values
E <- nrow(d01)/4
fls_E <- sample(c(E+1, E-1),1)

X2 <- round(test$statistic,4)
fls_X2 <- sample(c(X2-runif(1), X2+runif(1)),1)

p <- round(test$p.value, 4)
fls_p <- sample(c(p-.01, p+.01),1)

X2crit <- abs(round(qchisq(.05, test$parameter, lower.tail=F),4))
fls_X2crit <- sample(c(X2crit-runif(1, min=.1), X2crit+runif(1, min=.1)),1)

df=3

# Question 1 (Expected Value)
if(runif(1) > 0.5) {
  questions[1] <- paste0("The expected value is **less** than $",fmt(fls_E,3),"$.")
  solutions[1] <- E < fls_E
  explanations[1] <- paste0("The correct answer is ",solutions[1],". The expected value is $",fmt(E,3),"$, and hence, ",ifelse(solutions[1], "**is**", "is **not**")," less than $",fmt(fls_E,3),"$.")
} else {
  questions[1] <- paste0("The expected value is **greater** than $",fmt(fls_E,3),"$.")
  solutions[1] <- E > fls_E
  explanations[1] <- paste0("The correct answer is ",solutions[1],". The expected value is $",fmt(E,3),"$, and hence, ",ifelse(solutions[1], "**is**", "is **not**")," greater than $",fmt(fls_E,3),"$.")
}

# Question 2 (obtained X2)
if(runif(1) > 0.5) {
    questions[2] <- paste0("The obtained value for *chi-square* is **less** than $",fmt(fls_X2,3),"$.")
    solutions[2] <- X2 < fls_X2
    explanations[2] <- paste0("The correct answer is ",solutions[2],". The obtained value for *chi-square* is $",fmt(X2,3),"$, and hence, ",ifelse(solutions[2], "**is**", "is **not**")," less than $",fmt(fls_X2,3),"$.")
  } else {
    questions[2] <- paste0("The obtained value for *chi-square* is **greater** than $",fmt(fls_X2,3),"$.")
    solutions[2] <- X2 > fls_X2
    explanations[2] <- paste0("The correct answer is ",solutions[2],". The obtained value for *chi-square* is $",fmt(X2,3),"$, and hence, ",ifelse(solutions[2], "**is**", "is **not**")," greater than $",fmt(fls_X2,3),"$.")
  }

# Question 3 (critical value for X2)
if(runif(1) > 0.5) {
    questions[3] <- paste0("The critical value (alpha = .05) for *chi-square* is **less** than $",fmt(fls_X2crit,3),"$.")
    solutions[3] <- X2crit < fls_X2crit
    explanations[3] <- paste0("The correct answer is ",solutions[3],". The critical value (alpha = .05) for *chi-square* with $",fmt(df,0),"$ degrees of freedom is $",fmt(X2crit,3),"$, and hence, ",ifelse(solutions[3], "**is**", "is **not**")," less than $",fmt(fls_X2crit,3),"$.")
  } else {
    questions[3] <- paste0("The critical value (alpha = .05) for *chi-square* is **greater** than $",fmt(fls_X2crit,3),"$.")
    solutions[3] <- X2crit > fls_X2crit
    explanations[3] <- paste0("The correct answer is ",solutions[3],". The critical value (alpha = .05) for *chi-square* with $",fmt(df,0),"$ degrees of freedom is $",fmt(X2crit,3),"$, and hence, ",ifelse(solutions[3], "**is**", "is **not**")," greater than $",fmt(fls_X2crit,3),"$.")
  }

# Question 4 (significance)
questions[4] <- paste0("The results of the *chi-square* test are statistically significant.")
solutions[4] <- X2 >= X2crit
explanations[4] <- paste0("The correct answer is ",solutions[4],". The  obtained value for *chi-square* is $",fmt(X2,3),"$, and hence, ",ifelse(solutions[4], "**is**", "is **not**")," greater than or equal to the critical value, $",fmt(X2crit,3),"$.")

```

```{r IfAllFalsesMakeOneTRUE}
#IF ALL FALSE, MAKE ONE TRUE

i <- 0

if(all(solutions==FALSE)) {
  i <- sample.int(4,1)
}

if(i==1) {
  solutions[1] <- TRUE
  if(E < fls_E) {
  questions[1] <- paste0("The expected value is **less** than $",fmt(fls_E,3),"$.")
  explanations[1] <- paste0("The correct answer is TRUE. The expected value is $",fmt(E,3),"$, and hence, **is** less than $",fmt(fls_E,3),"$.")
  } else {
    questions[1] <- paste0("The expected value is **greater** than $",fmt(fls_E,3),"$.")
    explanations[1] <- paste0("The correct answer is TRUE. The expected value is $",fmt(E,3),"$, and hence, **is** greater than $",fmt(fls_E,3),"$.")
  }
} else if (i==2) {
    solutions[2] <- TRUE
    if(X2 < fls_X2) {
    questions[2] <- paste0("The obtained value for *chi-square* is **less** than $",fmt(fls_X2,3),"$.")
    explanations[2] <- paste0("The correct answer is TRUE. The obtained value for *chi-square* is $",fmt(X2,3),"$, and hence, **is** less than $",fmt(fls_X2,3),"$.")
  } else {
    questions[2] <- paste0("The obtained value for *chi-square* is **greater** than $",fmt(fls_X2,3),"$.")
    explanations[2] <- paste0("The correct answer is TRUE. The obtained value for *chi-square* is $",fmt(X2,3),"$, and hence, **is** greater than $",fmt(fls_X2,3),"$.")
  }
} else if (i==3) {
    solutions[3] <- TRUE
    if(X2crit < fls_X2crit) {
    questions[3] <- paste0("The critical value (alpha = .05) for *chi-square* is **less** than $",fmt(fls_X2crit,3),"$.")
    explanations[3] <- paste0("The correct answer is TRUE. The critical value (alpha = .05) for *chi-square* with $",fmt(df,0),"$ degrees of freedom is $",fmt(X2crit,3),"$, and hence, **is** less than $",fmt(fls_X2crit,3),"$.")
  } else {
    questions[3] <- paste0("The critical value (alpha = .05) for *chi-square* is **greater** than $",fmt(fls_X2crit,3),"$.")
    explanations[3] <- paste0("The correct answer is TRUE. The critical value (alpha = .05) for *chi-square* with $",fmt(df,0),"$ degrees of freedom is $",fmt(X2crit,3),"$, and hence, **is** greater than $",fmt(fls_X2crit,3),"$.")
  }
} else if (i==4) {
    solutions[4] <- TRUE
    if(X2 >= X2crit) {
    questions[4] <- paste0("The results of the *chi-square* test are statistically significant.")
    explanations[4] <- paste0("The correct answer is TRUE. The obtained value for *chi-square* is $",fmt(X2,3),"$, and hence, **is** greater than or equal to the critical value, $",fmt(X2crit,3),"$.")
  } else {
    questions[4] <- paste0("The results of the *chi-square* test are **not** statistically significant.")
    explanations[4] <- paste0("The correct answer is TRUE. The obtained value for *chi-square* is $",fmt(X2,3),"$, and hence, is **not** greater than or equal to the critical value, $",fmt(X2crit,3),"$.")
  }
}

rm(i)

```

```{r IfAllTruesMakeOneFALSE}
j <- 0

if(all(solutions==TRUE)) {
  j <- sample.int(4,1)
}

if(j==1) {
  solutions[1] <- FALSE
  if(E > fls_E) {
  questions[1] <- paste0("The expected value is **less** than $",fmt(fls_E,3),"$.")
  explanations[1] <- paste0("The correct answer is FALSE. The expected value is $",fmt(E,3),"$, and hence, is **not** less than $",fmt(fls_E,3),"$.")
  } else {
    questions[1] <- paste0("The expected value is **greater** than $",fmt(fls_E,3),"$.")
    explanations[1] <- paste0("The correct answer is FALSE. The expected value is $",fmt(E,3),"$, and hence, is **not** greater than $",fmt(fls_E,3),"$.")
  }
} else if (j==2) {
    solutions[2] <- FALSE
    if(X2 > fls_X2) {
    questions[2] <- paste0("The obtained value for *chi-square* is **less** than $",fmt(fls_X2,3),"$.")
    explanations[2] <- paste0("The correct answer is FALSE. The obtained value for *chi-square* is $",fmt(X2,3),"$, and hence, is **not** less than $",fmt(fls_X2,3),"$.")
  } else {
    questions[2] <- paste0("The obtained value for *chi-square* is **greater** than $",fmt(fls_X2,3),"$.")
    explanations[2] <- paste0("The correct answer is FALSE. The obtained value for *chi-square* is $",fmt(X2,3),"$, and hence, is **not** greater than $",fmt(fls_X2,3),"$.")
  }
} else if (j==3) {
    solutions[3] <- FALSE
    if(X2crit > fls_X2crit) {
    questions[3] <- paste0("The critical value (alpha = .05) for *chi-square* is **less** than $",fmt(fls_X2crit,3),"$.")
    explanations[3] <- paste0("The correct answer is FALSE. The critical value (alpha = .05) for *chi-square* with $",fmt(df,0),"$ degrees of freedom is $",fmt(X2crit,3),"$, and hence, is **not** less than $",fmt(fls_X2crit,3),"$.")
  } else {
    questions[3] <- paste0("The critical value (alpha = .05) for *chi-square* is **greater** than $",fmt(fls_X2crit,3),"$.")
    explanations[3] <- paste0("The correct answer is FALSE. The critical value (alpha = .05) for *chi-square* with $",fmt(df,0),"$ degrees of freedom is $",fmt(X2crit,3),"$, and hence, is **not** greater than $",fmt(fls_X2crit,3),"$.")
  }
} else if (j==4) {
    solutions[4] <- FALSE
    if(X2 <= X2crit) {
    questions[4] <- paste0("The results of the *chi-square* test are statistically significant.")
    explanations[4] <- paste0("The correct answer is FALSE. The obtained value for *chi-square* is $",fmt(X2,3),"$, and hence, is **not** greater than or equal to the critical value, $",fmt(X2crit,3),"$.")
  } else {
    questions[4] <- paste0("The results of the *chi-square* test are **not** statistically significant.")
    explanations[4] <- paste0("The correct answer is FALSE. The obtained value for *chi-square* is $",fmt(X2,3),"$, and hence, **is** greater than or equal to the critical value, $",fmt(X2crit,3),"$.")
  }
}

rm(j)

```

```{r QuestionList, echo=F, results = "asis"}

answerlist(questions, markup = "markdown")

```

# Solution

The results of this *chi-square* test are as follows:

$$\chi^2(`r test$parameter`)=`r round(test$statistic,3)`, p=`r round(test$p.value,3)`$$

------------------------------------------------------------------------

The degrees of freedom for a a chi-square with $k$ categories is $k-1$. Thus, for the current data set it is as follows:

$$k-1=`r 4`-1=`r 4-1`$$

------------------------------------------------------------------------

The obtained value for the $\chi^2$ goodness-of-fit test is calculated with the following formula:

$$\chi^2=\sum\limits_{i=1}^k\frac{(O_i-E_i)^2}{E_i}$$

------------------------------------------------------------------------

The way to calculate this using the method covered in class is below:

|                  $Category$                  | $Observed$ |                   $Expected$                    |                               $O-E$                                |                      $(O-E)^2$                      |                                                                $(O-E)^2\big/E$                                                                 |
|:----------:|:----------:|:----------:|:----------:|:----------:|:----------:|
| $`r s02$ClassLevel[1]`$ | $`r s02$Count[1]`$ |  $`r E`$   | $`r s02$Count[1]-E`$ | $`r (s02$Count[1]-E)^2`$ | $`r round(((s02$Count[1]-E)^2)/E,4)`$ | | $`r s02$ClassLevel[4]`$ | $`r s02$Count[4]`$ | $`r E`$ | $`r s02$Count[4]-E`$ |                                       $`r (s02$Count[4]-E)^2`$ | $`r round(((s02$Count[4]-E)^2)/E,4)`$                                        |
| $`r s02$ClassLevel[2]`$ | $`r s02$Count[2]`$ |  $`r E`$   | $`r s02$Count[2]-E`$ | $`r (s02$Count[2]-E)^2`$ | $`r round(((s02$Count[2]-E)^2)/E,4)`$ | | $`r s02$ClassLevel[3]`$ | $`r s02$Count[3]`$ | $`r E`$ | $`r s02$Count[3]-E`$ |                                       $`r (s02$Count[3]-E)^2`$ | $`r round(((s02$Count[3]-E)^2)/E,4)`$                                        |
|                                              |            |                                                 |                                                                    |                                                     | $\sum$ $`r round(sum((((test$observed[1]-E)^2)/E),(((test$observed[4]-E)^2)/E),(((test$observed[2]-E)^2)/E),(((test$observed[3]-E)^2)/E)),3)`$ |

------------------------------------------------------------------------

To get the critical value for *chi-square* with $\alpha=.05$, you go to <https://patrickabolger.github.io/PSYC-301-Exam-Formulas/#chi2-table> and make the following choices:

-   *Degrees of freedom*
    -   $`r test$parameter`$
-   *Alpha level*
    -   0.05
-   Then look at the intersection of those two values for the critical value

------------------------------------------------------------------------

In case you are interested, a bar chart of this data is below:

```{r}
d01 %>% ggplot(aes(`Class Level`)) +
  geom_bar(fill="#500000")
```

<br/> Or a pie chart with percentages (out of a total of `r N`) instead of counts:

```{r}
s03 %>% ggplot(aes(x = "", y = percentage, fill = `Class Level`))+
  geom_bar(stat = "identity", color="white")+
  coord_polar("y", start = 200) +
  geom_text(aes(y = lab.pos, label = paste(percentage,"%", sep = "")), col = "white") +
  theme_void() +
  scale_fill_brewer(palette = "Set2")
```

```{r SolutionList, echo=F, results = "asis"}
answerlist(explanations, markup = "markdown")
```

# Meta-information

extype: mchoice exsolution: `r mchoice2string(solutions)` exname: X2GoF
