---
title: "correlations"
output: html_document
editor_options:
  markdown:
    canonical: false
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, message = NA)
```

```{r LoadPackages, include=F}

# install the packages below if needed
pkgs <- c("truncnorm", "tibble", "dplyr", "magrittr", "exams", "kableExtra", "ggplot2")
lapply(pkgs, library, character.only=T)

```

```{r SuppressScientificNotation}
options(scipen=999)
```

```{r data generation, echo = FALSE, message=F, warning=F, results = "hide"}

N <- sample(10:12,1) # random number between 10 and 12
Cons <- round(rtruncnorm(N, 1, 5, 4.17, 0.53),2)
error <- rnorm(N, 0, 30) # down from 104 in original (p. 472)
# This allows me to keep the same regression coefficient

FICO <- 694 + # 694 in original study (table 1, p. 473)
  45.97*Cons + # from table 2, p. 474 
  error

# Remove decimal places for FICO
FICO <- round(FICO,0)

# Put into a data frame for display

d <- as_tibble(cbind(FICO, Cons))

# For formula 1 with z-scores
meanFICO <- round(mean(FICO),4)
sdFICO <- round(sd(FICO),4)
zFICO <- round(((FICO-meanFICO)/(sdFICO)),4)

meanCons <- round(mean(Cons),4)
sdCons <- round(sd(Cons),4)
zCons <- round(((Cons-meanCons)/(sdCons)),4)

zFzC <- round(zFICO*zCons,4)
rz <- round(sum(zFzC)/(N-1),3)

# For formula 2 with raw scores
X <- round(FICO,4)
Y <- round(Cons,4)
Xsq <- round(X^2,4)
Ysq <- round(Y^2,4)
XY <- round(X*Y,4)

rlong_num <- N*(sum(XY))-(sum(X)*sum(Y))
rlong_den <- sqrt(((N*(sum(Xsq)))-(sum(X))^2)*((N*(sum(Ysq)))-(sum(Y))^2))
rlong <- round(rlong_num/rlong_den,3)

# Officially from the cor.test function

test <- cor.test(FICO, Cons)

```

# Question

Bernerth, Taylor, Walker, and Whitman (2012) studied the effect of the
*Big Five* personality traits (*Neuroticism*, *Conscientiousness*,
*Extraversion*, *Agreeableness*, and *Openness*; Likert scale from 1-5)
on FICO ("credit rating") credit scores (ranging from 300 - 850) in an
attempt to elucidate to employers what FICO scores may or may not
suggest about people applying for employment. Among other things, they
found that, with a group of 113 participants, FICO scores (*M* = 694.45,
*SD* = 104.22) correlated significantly with the *Big-5* personality
trait *Conscientiousness* (*M* = 4.17, *SD* = 0.53). Specifically, the
two variables correlated with a Pearson's *r* of .20, a positive
correlation: *r*(111) = .20, *p* \< .05.

You want to replicate this portion of the study with a small number of
participants, so you recruit `r N` participants who are willing to take
the same *Big-5* personality-trait survey, as well as provide you their
FICO scores.

------------------------------------------------------------------------

**Calculations**

The raw scores of your `r N` participants are below. For the FICO scores
(some of which may go beyond the maximum possible), the mean is
`r round(meanFICO,4)` (*SD* = `r round(sdFICO,4)`)). For the Big-5
characteristic *Conscientiousness*, the mean is `r round(meanCons,4)`
(*SD* = `r round(sdCons,4)`). Calculate Pearson's *r* for the
association between these two variables. Round to ***FOUR*** decimal
places at every step, with your final answer rounded to ***THREE***
decimal places.

```{r DisplayRawScores, echo = FALSE, message=F, warning=F, fig.cap=''}
d %>% 
  kable(digits=c(3), align = "c") %>%
  kable_styling("striped", full_width = FALSE)
```

------------------------------------------------------------------------

<br/>

**TRUE/FALSE Questions**

Now that you have calculated Pearson's *r*, answer the questions below.
Check the box to indicate **TRUE**, and leave it blank to indicate
**FALSE**.

> To answer all the questions, you will need to refer to this [Pearson's
> r lookup
> table](https://patrickabolger.github.io/PSYC-301-Exam-Formulas/#table-for-pearsons-r)

------------------------------------------------------------------------

<br/>

```{r QuestionVectorSize, echo=F, results = "hide"}
questions <- character(4)
solutions <- logical(4)
explanations <- character(4)
```

```{r QuestionGeneration, echo=F, results = "hide"}
# Create a false df, t, and p value above or below the true values
p <- round(test$p.value,3)
df <- test$parameter
r <- round(test$estimate,3)
abs_r <- abs(r)
rcrit <- ifelse(df==8,.632, 
                ifelse(df==9,.602,
                       ifelse(df==10,.576,
                                            NA)))
# ^^ specific to the N setting, hard coded
# 
fls_df <- sample(c(N-4,N-3,N-1,N,N+1,N+2,N+3), 1)
fls_r <- round(sample(c(r - .005*r, r + .005*r),1), 3)
fls_rcrit <- round(sample(c(rcrit - .01*rcrit, rcrit + .01*rcrit),1), 3)

# Question 1 (df)
if(runif(1) > 0.5) {
  questions[1] <- paste0("The **degrees of freedom** are **less** than $",fmt(fls_df,0),"$.")
  solutions[1] <- df < fls_df
  explanations[1] <- paste0("The correct answer is ",solutions[1],". The degrees of freedom are $",fmt(df,0),"$, and hence, ",ifelse(solutions[1], "**are**", "are **not**")," less than $",fmt(fls_df,0),"$.")
} else {
  questions[1] <- paste0("The **degrees of freedom** are **greater** than $",fmt(fls_df,0),"$.")
  solutions[1] <- df > fls_df
  explanations[1] <- paste0("The correct answer is ",solutions[1],". The degrees of freedom are $",fmt(df,0),"$, and hence, ",ifelse(solutions[1], "**are**", "are **not**")," greater than $",fmt(fls_df,0),"$.")
}

# Question 2 (obtained r)
if(runif(1) > 0.5) {
    questions[2] <- paste0("The obtained value for *r* is **less** than $",fmt(fls_r,3),"$.")
    solutions[2] <- r < fls_r
    explanations[2] <- paste0("The correct answer is ",solutions[2],". The obtained value for *r* is $",fmt(r,3),"$, and hence, ",ifelse(solutions[2], "**is**", "is **not**")," less than $",fmt(fls_r,3),"$.")
  } else {
    questions[2] <- paste0("The obtained value for *r* is **greater** than $",fmt(fls_r,3),"$.")
    solutions[2] <- r > fls_r
    explanations[2] <- paste0("The correct answer is ",solutions[2],". The obtained value for *r* is $",fmt(r,3),"$, and hence, ",ifelse(solutions[2], "**is**", "is **not**")," greater than $",fmt(fls_r,3),"$.")
  }

# Question 3 (critical value for r)
if(runif(1) > 0.5) {
    questions[3] <- paste0("The critical value (alpha = .05) for *r* is **less** than $",fmt(fls_rcrit,3),"$.")
    solutions[3] <- rcrit < fls_rcrit
    explanations[3] <- paste0("The correct answer is ",solutions[4],". The critical value (alpha = .05) for *r* with $",fmt(df,0),"$ degrees of freedom is $",fmt(rcrit,3),"$, and hence, ",ifelse(solutions[3], "**is**", "is **not**")," less than $",fmt(fls_rcrit,3),"$.")
  } else {
    questions[3] <- paste0("The critical value (alpha = .05) for *r* is **greater** than $",fmt(fls_rcrit,3),"$.")
    solutions[3] <- rcrit > fls_rcrit
    explanations[3] <- paste0("The correct answer is ",solutions[3],". The critical value (alpha = .05) for *r* with $",fmt(df,0),"$ degrees of freedom is $",fmt(rcrit,3),"$, and hence, ",ifelse(solutions[3], "**is**", "is **not**")," greater than $",fmt(fls_rcrit,3),"$.")
  }

# Question 4 (significance)
questions[4] <- paste0("The results of the correlation (Pearson's *r*) are statistically significant.")
solutions[4] <- abs_r >= rcrit
explanations[4] <- paste0("The correct answer is ",solutions[4],". The absolute value of the obtained value for *r* is $",fmt(abs_r,3),"$, and hence, ",ifelse(solutions[4], "**is**", "is **not**")," greater than or equal to the critical value, $",fmt(rcrit,3),"$.")

```

```{r ifAllFalseMakeOneTRUE}
i <- 0
if(all(solutions==FALSE)) {
i <- sample.int(4,1)
}

if(i==1) {
  solutions[1] <- TRUE
  if(df < fls_df) {
  questions[1] <- paste0("The **degrees of freedom** are **less** than $",fmt(fls_df,0),"$.")
  explanations[1] <- paste0("The correct answer is TRUE. The degrees of freedom are $",fmt(df,0),"$, and hence, are less than $",fmt(fls_df,0),"$.")
  } else {
    questions[1] <- paste0("The **degrees of freedom** are **greater** than $",fmt(fls_df,0),"$.")
    explanations[1] <- paste0("The correct answer is TRUE. The degrees of freedom are $",fmt(df,0),"$, and hence, are greater than $",fmt(fls_df,0),"$.")
  }
} else if (i==2) {
    solutions[2] <- TRUE
    if(r < fls_r) {
    questions[2] <- paste0("The obtained value for *r* is **less** than $",fmt(fls_r,3),"$.")
    explanations[2] <- paste0("The correct answer is TRUE. The obtained value for *r* is $",fmt(r,3),"$, and hence, is less than $",fmt(fls_r,3),"$.")
  } else {
    questions[2] <- paste0("The obtained value for *r* is **greater** than $",fmt(fls_r,3),"$.")
    explanations[2] <- paste0("The correct answer is TRUE. The obtained value for *r* is $",fmt(r,3),"$, and hence, is greater than $",fmt(fls_r,3),"$.")
  }
} else if (i==3) {
    solutions[3] <- TRUE
    if(rcrit < fls_rcrit) {
    questions[3] <- paste0("The critical value (alpha = .05) for *r* is **less** than $",fmt(fls_rcrit,3),"$.")
    explanations[3] <- paste0("The correct answer is TRUE. The critical value (alpha = .05) for *r* with $",fmt(df,0),"$ degrees of freedom is $",fmt(rcrit,3),"$, and hence, is less than $",fmt(fls_rcrit,3),"$.")
  } else {
    questions[3] <- paste0("The critical value (alpha = .05) for *r* is **greater** than $",fmt(fls_rcrit,3),"$.")
    explanations[3] <- paste0("The correct answer is TRUE. The critical value (alpha = .05) for *r* with $",fmt(df,0),"$ degrees of freedom is $",fmt(rcrit,3),"$, and hence, is greater than $",fmt(fls_rcrit,3),"$.")
  }
} else if (i==4) {
    solutions[4] <- TRUE
    if (abs_r >= rcrit) {
questions[4] <- paste0("The results of the correlation (Pearson's *r*) are statistically significant.")
explanations[4] <- paste0("The correct answer is TRUE. The absolute value of the obtained value for *r* is $",fmt(abs_r,3),"$, and hence, is greater than or equal to the critical value, $",fmt(rcrit,3),"$.")
  } else {
questions[4] <- paste0("The results of the correlation (Pearson's *r*) are **not** statistically significant.")
explanations[4] <- paste0("The correct answer is TRUE. The absolute value of the obtained value for *r* is $",fmt(abs_r,3),"$, and hence, is less than the critical value, $",fmt(rcrit,3),"$.")
  }
}

rm(i)

```

```{r ifAllTrueMakeOneFALSE}
j <- 0
if(all(solutions==TRUE)) {
j <- sample.int(4,1)
}

if(j==1) {
  solutions[1] <- FALSE
  if(df > fls_df) {
  questions[1] <- paste0("The **degrees of freedom** are **less** than $",fmt(fls_df,0),"$.")
  explanations[1] <- paste0("The correct answer is FALSE. The degrees of freedom are $",fmt(df,0),"$, and hence, are greater than $",fmt(fls_df,0),"$.")
  } else {
    questions[1] <- paste0("The **degrees of freedom** are **greater** than $",fmt(fls_df,0),"$.")
    explanations[1] <- paste0("The correct answer is FALSE. The degrees of freedom are $",fmt(df,0),"$, and hence, are less than $",fmt(fls_df,0),"$.")
  }
} else if (j==2) {
    solutions[2] <- FALSE
    if(r > fls_r) {
    questions[2] <- paste0("The obtained value for *r* is **less** than $",fmt(fls_r,3),"$.")
    explanations[2] <- paste0("The correct answer is FALSE. The obtained value for *r* is $",fmt(r,3),"$, and hence, is greater than $",fmt(fls_r,3),"$.")
  } else {
    questions[2] <- paste0("The obtained value for *r* is **greater** than $",fmt(fls_r,3),"$.")
    explanations[2] <- paste0("The correct answer is FALSE. The obtained value for *r* is $",fmt(r,3),"$, and hence, is less than $",fmt(fls_r,3),"$.")
  }
} else if (j==3) {
    solutions[3] <- FALSE
    if(rcrit > fls_rcrit) {
    questions[3] <- paste0("The critical value (alpha = .05) for *r* is **less** than $",fmt(fls_rcrit,3),"$.")
    explanations[3] <- paste0("The correct answer is FALSE. The critical value (alpha = .05) for *r* with $",fmt(df,0),"$ degrees of freedom is $",fmt(rcrit,3),"$, and hence, is greater than $",fmt(fls_rcrit,3),"$.")
  } else {
    questions[3] <- paste0("The critical value (alpha = .05) for *r* is **greater** than $",fmt(fls_rcrit,3),"$.")
    explanations[3] <- paste0("The correct answer is FALSE. The critical value (alpha = .05) for *r* with $",fmt(df,0),"$ degrees of freedom is $",fmt(rcrit,3),"$, and hence, is less than $",fmt(fls_rcrit,3),"$.")
  }
} else if (j==4) {
    solutions[4] <- FALSE
    if (abs_r < rcrit) {
questions[4] <- paste0("The results of the correlation (Pearson's *r*) are statistically significant.")
explanations[4] <- paste0("The correct answer is FALSE. The absolute value of the obtained value for *r* is $",fmt(abs_r,3),"$, and hence, is less than the critical value, $",fmt(rcrit,3),"$.")
  } else {
questions[4] <- paste0("The results of the correlation (Pearson's *r*) are **not** statistically significant.")
explanations[4] <- paste0("The correct answer is FALSE. The absolute value of the obtained value for *r* is $",fmt(abs_r,3),"$, and hence, is greater than or equal to the critical value, $",fmt(rcrit,3),"$.")
  }
}

rm(j)

```

```{r QuestionList, echo=F, results = "asis"}
answerlist(questions, markup = "markdown")
```

# Solution

The degrees of freedom for a Pearson's *r* is $N-2$. Thus, for the
current data set it is $N-2=`r N`-2=`r df`$.

<br/>

The obtained value for Pearson's *r* is calculated with any of the
following formulas.

Using *z*-scores:

## $$r=\frac{\sum{z_{x_i}z_{y_i}}}{n-1}$$

Using raw scores:

$$r=\frac{n\sum{XY}-\sum{X}\sum{Y}}{\sqrt{\Big[n\sum{X^2-\big(\sum{X}\big)^2}\Big]\Big[n\sum{Y^2}-\big(\sum{Y}\big)^2\Big]}}$$

(***NOTE***: There may be slight rounding-error differences between the
two approaches above. But in bench-testing, this difference was never
more than a .001 difference. This minuscule difference is not enough to
affect any TRUE/FALSE answers.)

------------------------------------------------------------------------

Here is the solution using *z*-scores:

|               $x_i$                |             $\bar{x}$              |              $s_x$               |              $z_{x_i}$              |          $z_{x_i}z_{y_i}$          |              $z_{y_i}$              |              $s_y$               |             $\bar{y}$              |               $y_i$                |
|:------:|:------:|:------:|:------:|:------:|:------:|:------:|:------:|:------:|
|            `r FICO[1]`             |            `r meanFICO`            |            `r sdFICO`            |            `r zFICO[1]`             |            `r zFzC[1]`             |            `r zCons[1]`             |            `r sdCons`            |            `r meanCons`            |            `r Cons[1]`             |
|            `r FICO[2]`             |            `r meanFICO`            |            `r sdFICO`            |            `r zFICO[2]`             |            `r zFzC[2]`             |            `r zCons[2]`             |            `r sdCons`            |            `r meanCons`            |            `r Cons[2]`             |
|            `r FICO[3]`             |            `r meanFICO`            |            `r sdFICO`            |            `r zFICO[3]`             |            `r zFzC[3]`             |            `r zCons[3]`             |            `r sdCons`            |            `r meanCons`            |            `r Cons[3]`             |
|            `r FICO[4]`             |            `r meanFICO`            |            `r sdFICO`            |            `r zFICO[4]`             |            `r zFzC[4]`             |            `r zCons[4]`             |            `r sdCons`            |            `r meanCons`            |            `r Cons[4]`             |
|            `r FICO[5]`             |            `r meanFICO`            |            `r sdFICO`            |            `r zFICO[5]`             |            `r zFzC[5]`             |            `r zCons[5]`             |            `r sdCons`            |            `r meanCons`            |            `r Cons[5]`             |
|            `r FICO[6]`             |            `r meanFICO`            |            `r sdFICO`            |            `r zFICO[6]`             |            `r zFzC[6]`             |            `r zCons[6]`             |            `r sdCons`            |            `r meanCons`            |            `r Cons[6]`             |
|            `r FICO[7]`             |            `r meanFICO`            |            `r sdFICO`            |            `r zFICO[7]`             |            `r zFzC[7]`             |            `r zCons[7]`             |            `r sdCons`            |            `r meanCons`            |            `r Cons[7]`             |
|            `r FICO[8]`             |            `r meanFICO`            |            `r sdFICO`            |            `r zFICO[8]`             |            `r zFzC[8]`             |            `r zCons[8]`             |            `r sdCons`            |            `r meanCons`            |            `r Cons[8]`             |
|            `r FICO[9]`             |            `r meanFICO`            |            `r sdFICO`            |            `r zFICO[9]`             |            `r zFzC[9]`             |            `r zCons[9]`             |            `r sdCons`            |            `r meanCons`            |            `r Cons[9]`             |
|            `r FICO[10]`            |            `r meanFICO`            |            `r sdFICO`            |            `r zFICO[10]`            |            `r zFzC[10]`            |            `r zCons[10]`            |            `r sdCons`            |            `r meanCons`            |            `r Cons[10]`            |
| `r if (N>10) {FICO[11]} else {""}` | `r if (N>10) {meanFICO} else {""}` | `r if (N>10) {sdFICO} else {""}` | `r if (N>10) {zFICO[11]} else {""}` | `r if (N>10) {zFzC[11]} else {""}` | `r if (N>10) {zCons[11]} else {""}` | `r if (N>10) {sdCons} else {""}` | `r if (N>10) {meanCons} else {""}` | `r if (N>10) {Cons[11]} else {""}` |
| `r if (N>11) {FICO[12]} else {""}` | `r if (N>11) {meanFICO} else {""}` | `r if (N>11) {sdFICO} else {""}` | `r if (N>11) {zFICO[12]} else {""}` | `r if (N>11) {zFzC[12]} else {""}` | `r if (N>11) {zCons[12]} else {""}` | `r if (N>11) {sdCons} else {""}` | `r if (N>11) {meanCons} else {""}` | `r if (N>11) {Cons[12]} else {""}` |
|                                    |                                    |                                  |                                     |   $\sum$ `r round(sum(zFzC),4)`    |                                     |                                  |                                    |                                    |

Now you just need to divide this number by $n-1$:

$$r_{x,y}=\frac{\sum{z_{x_i}z_{y_i}}}{n-1}$$ <br/>
$$=\frac{`r round(sum(zFzC),4)`}{`r N`-1}$$ <br/>
$$=\frac{`r round(sum(zFzC),4)`}{`r N-1`}$$ <br/>
$$=`r round((sum(zFzC)/(N-1)),3)`$$ <br/>

------------------------------------------------------------------------

Here is the solution using raw scores

|               $X$               |              $Y$               |              $X^2$               |              $Y^2$               |              $XY$               |
|:-------------:|:-------------:|:-------------:|:-------------:|:-------------:|
|            `r X[1]`             |            `r Y[1]`            |            `r Xsq[1]`            |            `r Ysq[1]`            |            `r XY[1]`            |
|            `r X[2]`             |            `r Y[2]`            |            `r Xsq[2]`            |            `r Ysq[2]`            |            `r XY[2]`            |
|            `r X[3]`             |            `r Y[3]`            |            `r Xsq[3]`            |            `r Ysq[3]`            |            `r XY[3]`            |
|            `r X[4]`             |            `r Y[4]`            |            `r Xsq[4]`            |            `r Ysq[4]`            |            `r XY[4]`            |
|            `r X[5]`             |            `r Y[5]`            |            `r Xsq[5]`            |            `r Ysq[5]`            |            `r XY[5]`            |
|            `r X[6]`             |            `r Y[6]`            |            `r Xsq[6]`            |            `r Ysq[6]`            |            `r XY[6]`            |
|            `r X[7]`             |            `r Y[7]`            |            `r Xsq[7]`            |            `r Ysq[7]`            |            `r XY[7]`            |
|            `r X[8]`             |            `r Y[8]`            |            `r Xsq[8]`            |            `r Ysq[8]`            |            `r XY[8]`            |
|            `r X[9]`             |            `r Y[9]`            |            `r Xsq[9]`            |            `r Ysq[9]`            |            `r XY[9]`            |
|            `r X[10]`            |           `r Y[10]`            |           `r Xsq[10]`            |           `r Ysq[10]`            |         `r X[10]*Y[10]`         |
| `r if (N>10) {X[11]} else{""}`  | `r if (N>10) {Y[11]} else{""}` | `r if (N>10) {Xsq[11]} else{""}` | `r if (N>10) {Ysq[11]} else{""}` | `r if (N>10) {XY[11]} else{""}` |
| `r if (N>11) {X[12]} else {""}` | `r if (N>11) {Y[12]} else{""}` | `r if (N>11) {Xsq[12]} else{""}` | `r if (N>11) {Ysq[12]} else{""}` | `r if (N>11) {XY[12]} else{""}` |
|   $\sum$ `r round(sum(X),3)`    |   $\sum$ `r round(sum(Y),3)`   |   $\sum$ `r round(sum(Xsq),3)`   |   $\sum$ `r round(sum(Ysq),3)`   |   $\sum$ `r round(sum(XY),3)`   |

<br/> Plugging these numbers into the following equation,...

$$r=\frac{n\sum{XY}-\sum{X}\sum{Y}}{\sqrt{\Big[n\sum{X^2-\big(\sum{X}\big)^2}\Big]\Big[n\sum{Y^2}-\big(\sum{Y}\big)^2\Big]}}$$
<br/> ...we get the following: <br/>

$$r=\frac{`r N`(`r sum(XY)`)-(`r sum(X)`)(`r sum(Y)`)}{\sqrt{\Big[`r N`(`r sum(Xsq)`)-\big(`r sum(X)`\big)^2\Big]\Big[`r N`(`r sum(Ysq)`)-\big(`r sum(Y)`\big)^2\Big]}}$$
<br/>

$$r=\frac{`r round(N*sum(XY),4)`-`r round(sum(X)*sum(Y),4)`}{\sqrt{\Big[`r round(N*sum(Xsq),4)`-`r round(sum(X)^2,4)`\Big]\Big[`r round(N*sum(Ysq),4)`-`r round(sum(Y)^2,4)`\Big]}}$$
<br/>

$$r=\frac{`r round(N*sum(XY)-sum(X)*sum(Y),4)`}{\sqrt{\Big[`r round(N*sum(Xsq)-sum(X)^2,4)`\Big]\Big[`r round(N*sum(Ysq)-sum(Y)^2,4)`\Big]}}$$
<br/>

$$r=\frac{`r round(N*sum(XY)-sum(X)*sum(Y),4)`}{\sqrt{`r round((N*sum(Xsq)-sum(X)^2)*(N*sum(Ysq)-sum(Y)^2),4)`}}$$
<br/>

$$r=\frac{`r round(N*sum(XY)-sum(X)*sum(Y),4)`}{`r round(sqrt( (N*sum(Xsq)-sum(X)^2)*(N*sum(Ysq)-sum(Y)^2)),4)`}$$
<br/>

$$r=`r round(((N*sum(XY))-(sum(X)*sum(Y)))/(sqrt((N*sum(Xsq)-sum(X)^2)*(N*sum(Ysq)-sum(Y)^2))),3)`$$

------------------------------------------------------------------------

To get the critical value (two-tailed) for *r* with $\alpha=.05$, you
need to refer to a critical-values table for *r* like here:
<https://patrickabolger.github.io/PSYC-301-Exam-Formulas/#table-for-pearsons-r>

Recall that the results of this correlation (Pearson's *r*) are as
follows:

$$r_{x,y}(`r round(df,0)`) = `r round(r,4)`,p=`r round(p,4)`$$

(though you weren't asked to calculate the exact *p*-value)

------------------------------------------------------------------------

On a scatterplot, the data would look like the following:

```{r, echo=F, warning=F, message=F}
d %>% ggplot(aes(scale(Cons), scale(FICO))) +
  geom_point(color="#500000") +
  geom_smooth(method = lm, color = "#500000")
```

------------------------------------------------------------------------

**Reference**

Bernerth, J. B., Taylor, S. G., Walker, H. J., Whitman, D. S. (2012). An
empirical investigation of dispositional antecedents and
performance-related outcomes of credit scores. *Journal of Applied
Psychology*, *97*(2), 469-478.

```{r SolutionList, echo=F, results = "asis"}
answerlist(explanations, markup = "markdown")
```

# Meta-information

extype: mchoice exsolution: `r mchoice2string(solutions)` exname:
correlation
