---
title: "linear regression"
output: html_document
editor_options:
  markdown:
    canonical: false
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, message = NA)
```

```{r LoadPackages, include=F}

# install the packages below if needed
pkgs <- c("truncnorm", "tibble", "dplyr", "magrittr", "exams", "kableExtra", "ggplot2", "sjPlot")
lapply(pkgs, library, character.only=T)

```

```{r SuppressScientificNotation}
options(scipen=999)
```

```{r data generation, echo = FALSE, message=F, warning=F, results = "hide"}

N <- sample(10:12,1) # random number between 10 and 13
Agreeableness <- round(rtruncnorm(N, 1, 5, 4.24, 0.53),2)
error <- rnorm(N, 0, 25) # down from 104 in original (p. 472)
# This allows me to keep the same regression coefficient

FICO <- 694 - # 694 in original study (table 1, p. 473)
  28.56*Agreeableness + # from table 2, p. 474 
  error

# Remove decimal places for FICO
FICO <- round(FICO,0)

d01 <- as_tibble(cbind(FICO, Agreeableness))

meanFICO <- round(mean(d01$FICO),3)
sdFICO <- round(sd(d01$FICO),3)
meanAgrb <- round(mean(d01$Agreeableness),3)
sdAgrb <- round(sd(d01$Agreeableness),3)

test <- lm(FICO ~ Agreeableness,data = d01)
s01 <- summary(test)
#s01 # to see results during development only
```

```{r solutionVars, echo = FALSE, message=F, warning=F, results = "hide"}
# Variables
X <- Agreeableness
Y <- FICO
Xsq <- Agreeableness^2
Ysq <- FICO^2
XY <- Agreeableness*FICO
# Slope and Intercept
b1 <- ((sum(XY))-(sum(X)*sum(Y)/N))/((sum(Xsq))-((sum(X)^2/N)))
b0 <- (sum(Y)-b1*sum(X))/N
```

# Question

Bernerth, Taylor, Walker, and Whitman (2012) found that FICO scores (*M*
= 694.45, *SD* = 104.22) correlated significantly with the personality
trait *Agreeableness* (*M* = 4.24, *SD* = 0.53). Specifically, for every
1-unit increase of *Agreeableness*, FICO scores ***dropped*** by 35.12
points, *b* = -35.12, *p* \< .05.

You want to replicate this portion of the study with a small number of
participants, so you recruit `r N` participants who are willing to take
the same *Big-5* personality-trait survey, as well as provide you their
FICO scores.

------------------------------------------------------------------------

<br/>

**Calculations**

The raw scores of your `r N` participants are below. For the FICO scores
(some of which may go beyond the maximum possible), the mean is
`r round(meanFICO,0)` (*SD* = `r round(sdFICO,3)`). For the Big 5
characteristic *Agreeableness*, the mean is `r round(meanAgrb,3)` (*SD*
= `r round(sdAgrb,3)`). For a simple linear regression, calculate both
the slope and the intercept for association between these two variables.
Round to ***FOUR*** decimal places at each step, but to ***THREE***
decimal places in your final answer.

You can refer to the following page for the formulas: [PSYC 301 exam
formulas](https://patrickabolger.github.io/PSYC-301-Exam-Formulas/#simple-regression)

```{r DisplayRawScores, echo = FALSE, message=F, warning=F, fig.cap=''}
d01 %>% 
  kable(digits=c(3), align = "c") %>%
  kable_styling("striped", full_width = FALSE)
```

------------------------------------------------------------------------

<br/>

**TRUE/FALSE Questions**

Now that you have calculated these statistics, answer the questions
below. Check the box to indicate **TRUE**, and leave it blank to
indicate **FALSE**.

------------------------------------------------------------------------

<br/>

```{r QuestionVectorSize, echo=F, results = "hide"}
questions <- character(2)
solutions <- logical(2)
explanations <- character(2)
```

```{r QuestionGeneration, echo=F, results = "hide"}
# Create a false df, t, and p value above or below the true values
int <- round(test$coefficients[1],3)
slp <- round(test$coefficients[2],3)

# ^^ specific to the N setting, hard coded
# 
fls_slp <- round(sample(c(slp - .005*slp, slp + .005*slp),1), 3)
fls_int <- round(sample(c(int - .005*int, int + .005*int),1), 3)

# Question 1 (slope)
if(runif(1) > 0.5) {
    questions[1] <- paste0("The slope, *b*, for `FICO` regressed on `Agreeableness` is **less** than $",fmt(fls_slp,3),"$.")
    solutions[1] <- slp < fls_slp
    explanations[1] <- paste0("The correct answer is ",solutions[1],". The obtained value for the slope for `FICO` regressed on `Agreeableness` is $",fmt(slp,3),"$, and hence, ",ifelse(solutions[1], "**is**", "is **not**")," less than $",fmt(fls_slp,3),"$.")
  } else {
    questions[1] <- paste0("The slope, *b*, for `FICO` regressed on `Agreeableness` is **greater** than $",fmt(fls_slp,3),"$.")
    solutions[1] <- slp > fls_slp
    explanations[1] <- paste0("The correct answer is ",solutions[1],". The slope for `FICO` regressed on `Agreeableness` is $",fmt(slp,3),"$, and hence, ",ifelse(solutions[1], "**is**", "is **not**")," greater than $",fmt(fls_slp,3),"$.")
  }

# Question 2 (intercept)
if(runif(1) > 0.5) {
    questions[2] <- paste0("The intercept, *a*, for `FICO` is **less** than $",fmt(fls_int,3),"$.")
    solutions[2] <- int < fls_int
    explanations[2] <- paste0("The correct answer is ",solutions[2],". The intercept for `FICO` is $",fmt(int,3),"$, and hence, ",ifelse(solutions[2], "**is**", "is **not**")," less than $",fmt(fls_int,3),"$.")
  } else {
    questions[2] <- paste0("The intercept, *a*, for `FICO` is **greater** than $",fmt(fls_int,3),"$.")
    solutions[2] <- int > fls_int
    explanations[2] <- paste0("The correct answer is ",solutions[2],". The intercept for `FICO` is $",fmt(int,3),"$, and hence, ",ifelse(solutions[2], "**is**", "is **not**")," greater than $",fmt(fls_int,3),"$.")
  }

```

```{r ifAllFalseMakeOneTRUE}

i <- 0

if(all(solutions==FALSE)) {
  i <- sample.int(2,1)
}

if(i==1) {
  solutions[1] <- TRUE
  if(slp < fls_slp) {
  questions[1] <- paste0("The slope, *b*, for `FICO` regressed on `Agreeableness` is **less** than $",fmt(fls_slp,3),"$.")
  explanations[1] <- paste0("The correct answer is TRUE. The slope for `FICO` regressed on `Agreeableness` is $",fmt(slp,3),"$, and hence, is less than $",fmt(fls_slp,3),"$.")
  } else {
    questions[1] <- paste0("The slope, *b*, for `FICO` regressed on `Agreeableness`is **greater** than $",fmt(fls_slp,3),"$.")
    explanations[1] <- paste0("The correct answer is TRUE. The slope for `FICO` regressed on `Agreeableness` is $",fmt(slp,3),"$, and hence, is greater than $",fmt(fls_slp,3),"$.")
  }
} else if (i==2) {
    solutions[2] <- TRUE
    if(int < fls_int) {
    questions[2] <- paste0("The intercept, *a*, for `FICO` is **less** than $",fmt(fls_int,3),"$.")
    explanations[2] <- paste0("The correct answer is TRUE. The intercept for `FICO` is $",fmt(int,3),"$, and hence, is less than $",fmt(fls_int,3),"$.")
  } else {
    questions[2] <- paste0("The intercept, *a*, for `FICO` is **greater** than $",fmt(fls_int,3),"$.")
    explanations[2] <- paste0("The correct answer is TRUE. The intercept for `FICO` is $",fmt(int,3),"$, and hence, is greater than $",fmt(fls_int,3),"$.")
  }
}

rm(i)

```

```{r ifAllTrueMakeOneFALSE}

j <- 0

if(all(solutions==TRUE)) {
  j <- sample.int(2,1)
}

if(j==1) {
  solutions[1] <- FALSE
  if(slp > fls_slp) {
  questions[1] <- paste0("The slope, *b*, for `FICO` regressed on `Agreeableness` is **less** than $",fmt(fls_slp,3),"$.")
  explanations[1] <- paste0("The correct answer is FALSE. The slope for `FICO` regressed on `Agreeableness` is $",fmt(slp,3),"$, and hence, is greater than $",fmt(fls_slp,3),"$.")
  } else {
    questions[1] <- paste0("The slope, *b*, for `FICO` regressed on `Agreeableness`is **greater** than $",fmt(fls_slp,3),"$.")
    explanations[1] <- paste0("The correct answer is FALSE. The slope for `FICO` regressed on `Agreeableness` is $",fmt(slp,3),"$, and hence, is less than $",fmt(fls_slp,3),"$.")
  }
} else if (j==2) {
    solutions[2] <- FALSE
    if(int > fls_int) {
    questions[2] <- paste0("The intercept, *a*, for `FICO` is **less** than $",fmt(fls_int,3),"$.")
    explanations[2] <- paste0("The correct answer is FALSE. The intercept for `FICO` is $",fmt(int,3),"$, and hence, is greater than $",fmt(fls_int,3),"$.")
  } else {
    questions[2] <- paste0("The intercept, *a*, for `FICO` is **greater** than $",fmt(fls_int,3),"$.")
    explanations[2] <- paste0("The correct answer is FALSE. The intercept for `FICO` is $",fmt(int,3),"$, and hence, is less than $",fmt(fls_int,3),"$.")
  }
}

rm(j)

```

```{r QuestionList, echo=F, results = "asis"}
answerlist(questions, markup = "markdown")
```

# Solution

The results of this simple regression are as follows:

```{r, echo=F, comment=NA}
tab_model(test, digits = 3)
```

------------------------------------------------------------------------

<br/>

The slope for a simple linear regression is calculated with the
following formula:

$$b=\frac{\sum{XY-(\sum{X}\sum{Y/n})}}{\sum{X}^2-[(\sum{X})^2/n]}$$
<br/>

So you need the following: $n$, $\sum{X}$, $\sum{Y}$, $\sum{X^2}$, and
finally $\sum{XY}$

The easiest way to do this is create a table and calculate all the
necessary components:

<br/>

|  $N$  |               $X$               |               $Y$               |               $X^2$               |               $XY$               |
|:-------------:|:-------------:|:-------------:|:-------------:|:-------------:|
|       |            `r X[1]`             |            `r Y[1]`             |            `r Xsq[1]`             |            `r XY[1]`             |
|       |            `r X[2]`             |            `r Y[2]`             |            `r Xsq[2]`             |            `r XY[2]`             |
|       |            `r X[3]`             |            `r Y[3]`             |            `r Xsq[3]`             |            `r XY[3]`             |
|       |            `r X[4]`             |            `r Y[4]`             |            `r Xsq[4]`             |            `r XY[4]`             |
|       |            `r X[5]`             |            `r Y[5]`             |            `r Xsq[5]`             |            `r XY[5]`             |
|       |            `r X[6]`             |            `r Y[6]`             |            `r Xsq[6]`             |            `r XY[6]`             |
|       |            `r X[7]`             |            `r Y[7]`             |            `r Xsq[7]`             |            `r XY[7]`             |
|       |            `r X[8]`             |            `r Y[8]`             |            `r Xsq[8]`             |            `r XY[8]`             |
|       |            `r X[9]`             |            `r Y[9]`             |            `r Xsq[9]`             |            `r XY[9]`             |
|       |            `r X[10]`            |            `r Y[10]`            |            `r Xsq[10]`            |            `r XY[10]`            |
|       | `r if (N>10) {X[11]} else {""}` | `r if (N>10) {Y[11]} else {""}` | `r if (N>10) {Xsq[11]} else {""}` | `r if (N>10) {XY[11]} else {""}` |
|       | `r if (N>11) {X[12]} else {""}` | `r if (N>11) {Y[12]} else {""}` | `r if (N>11) {Xsq[12]} else {""}` | `r if (N>11) {XY[12]} else {""}` |
| `r N` |   $\sum$ `r round(sum(X),4)`    |   $\sum$ `r round(sum(Y),4)`    |   $\sum$ `r round(sum(Xsq),4)`    |   $\sum$ `r round(sum(XY),4)`    |

<br/>

Once you have the sums (the bottom row) you put the sums into the
equation for the slope: <br/>
$$b=\frac{\sum{XY-(\sum{X}\sum{Y/n})}}{\sum{X}^2-[(\sum{X})^2/n]}$$

<br/>
$$b=\frac{`r round(sum(XY),4)`-[(`r round(sum(X),4)`)(`r round(sum(Y),4)`)/`r N`]}{`r round(sum(Xsq),4)`-[(`r round(sum(X),4)`^2)/`r N`]}$$
<br/>
$$b=\frac{`r round(sum(XY),4)`-[`r round(sum(X)*sum(Y),4)`/`r N`]}{`r round(sum(Xsq),4)`-[`r round((sum(X)^2),4)`/`r N`]}$$
<br/>
$$b=\frac{`r round(sum(XY),4)`-`r round((sum(X)*sum(Y))/N,4)`}{`r round(sum(Xsq),4)`-`r round((sum(X)^2)/N,4)`}$$
<br/>
$$b=\frac{`r round(sum(XY)-((sum(X)*sum(Y))/N),4)`}{`r round(sum(Xsq)-((sum(X)^2)/N),4)`}$$
<br/>
$$b=`r round((sum(XY)-((sum(X)*sum(Y))/N))/((sum(Xsq))-((sum(X)^2)/N)),3)`$$

------------------------------------------------------------------------

Once you have the slope, the intercept for a simple linear regression is
calculated with the following formula:

<br/> $$a=\frac{\sum{Y}-b\sum{X}}{n}$$ <br/> Thus:

<br/>
$$a=\frac{`r round(sum(Y),4)`-[(`r round(b1,4)`)(`r round(sum(X),4)`)]}{`r N`}$$
<br/> $$a=\frac{`r round(sum(Y),4)`-(`r round(b1*sum(X),4)`)}{`r N`}$$
<br/> $$a=\frac{`r round(sum(Y)-b1*sum(X),4)`}{`r N`}$$ <br/>
$$a=`r round((sum(Y)-b1*sum(X))/N,3)`$$

On a scatterplot, the data would look like the following:

```{r, echo=F, warning=F, message=F}
d01 %>% ggplot(aes(Agreeableness, FICO)) +
  geom_point(color="#500000") +
  geom_smooth(method = lm, color="#500000")
```

------------------------------------------------------------------------

***Reference***

Bernerth, J. B., Taylor, S. G., Walker, H. J., Whitman, D. S. (2012). An
empirical investigation of dispositional antecedents and
performance-related outcomes of credit scores. *Journal of Applied
Psychology*, *97*(2), 469-478.

```{r SolutionList, echo=F, results = "asis"}
answerlist(explanations, markup = "markdown")
```

# Meta-information

extype: mchoice exsolution: `r mchoice2string(solutions)` exname:
SimpleReg
