---
title: "Sums of Squares"
output: html_document
editor_options:
  markdown:
    canonical: false
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, message = NA)
```

```{r LoadPackages, include=F}

# install the packages below if needed
pkgs <- c("truncnorm", "tibble", "dplyr", "magrittr", "exams", "kableExtra", "ggplot2", "sjPlot", "car")
lapply(pkgs, library, character.only=T)

```

```{r SuppressScientificNotation}
options(scipen=999)
```

```{r dataGeneration, echo = FALSE, message=F, warning=F, results = "hide"}

Temperature <- c(
  rep("55", 3),
  rep("65", 3),
  rep("75", 3),
  rep("85", 3)
)

TestScore <- c(
  round(rnorm(3,72,6),0),
  round(rnorm(3,82,6),0),
  round(rnorm(3,85,6),0),
  round(rnorm(3,70,6),0)
)

d01 <- as.data.frame(cbind(Temperature, TestScore))
d01$Temperature <- as.factor(d01$Temperature)
d02 <- d01 %>% rename(`Test Score` = TestScore)
d02$`Test Score` <- as.numeric(d02$`Test Score`)
#rm(d01)
test <- anova(aov(`Test Score` ~ Temperature, data = d02))

```

```{r dataSolutions, echo = FALSE, message=F, warning=F, results = "hide"}

d <- d01 
d <- d %>% rename(x = Temperature, y = TestScore)
d$y <- round(as.numeric(d$y),4)
d$ybar <- round(mean(d$y),4)

d <- d %>%
  group_by(x) %>%
  mutate(ybark = mean(y)) %>%
  ungroup()

d$ybark <- round(d$ybark,4)

d$dev <- d$y-d$ybark
d$dev <- round(d$dev,4)

d$devSq <- d$dev^2
d$devSq <- round(d$devSq,4)

ssw <- sum(d$devSq)
ssw <- round(ssw,3)

#rm(d01)

dG <- d %>%
  group_by(x) %>%
  summarize(ybark=mean(y), n=n())

dG$ybark <- round(dG$ybark,4)

dG$ybar <- mean(d$y)
dG$ybar <- round(dG$ybar,4)

dG$dev <- dG$ybark - dG$ybar
dG$dev <- round(dG$dev,4)

dG$devSq <- dG$dev^2
dG$devSq <- round(dG$devSq,4)

dG$WdevSq <- dG$n*dG$devSq
dG$WdevSq <- round(dG$WdevSq,4)

ssb <- sum(dG$WdevSq)
ssb <- round(ssb,3)

```

# Question

You conducted a study in which four groups of randomly selected students
were given the same math exam (100 possible points) in rooms with
different mean temperatures: 85, 75, 65, and 55 degrees. Their scores on
the exams were recorded and analyzed by group. The data in the table
below are the test scores you recorded by group.

Your data are below:

------------------------------------------------------------------------

```{r DisplayRawScores, echo = FALSE, message=F, warning=F, fig.cap=''}
d02 %>% 
  kable(digits=c(3), align = "c") %>%
  pack_rows("55 Degrees", 1, 3) %>%
  pack_rows("65 Degrees", 4, 6) %>%
  pack_rows("75 Degrees", 7, 9) %>%
  pack_rows("85 Degrees", 10, 12) %>%
  kable_styling("striped", full_width = FALSE)
```

------------------------------------------------------------------------

<br/>

**Calculations**

In preparation for a One-way ANOVA, calculate both the *Sum of Squares
between* and the *Sum of Squares within*. Round to ***FOUR*** decimal
places at each step. Round the final answer to ***THREE*** decimal
places.

The formulas for the two types of *Sums of Squares* can be found on
[this
webpage](https://patrickabolger.github.io/PSYC-301-Exam-Formulas/#f-ratio-for-a-oneway-anova)

------------------------------------------------------------------------

**TRUE/FALSE Questions**

Now that you have calculated these statistics, answer the questions
below. Check the box to indicate **TRUE**, and leave it blank to
indicate **FALSE**.

------------------------------------------------------------------------

<br/>

```{r QuestionVectorSize, echo=F, results = "hide"}
questions <- character(2)
solutions <- logical(2)
explanations <- character(2)
```

```{r QuestionGeneration, echo=F, results = "hide"}
# Create a false df, t, and p value above or below the true values

fls_ssb <- round(sample(c(ssb - .005*ssb, ssb + .005*ssb),1), 3)
fls_ssw <- round(sample(c(ssw - .005*ssw, ssw + .005*ssw),1), 3)

# Question 1 (Sum of Squares Between)
if(runif(1) > 0.5) {
    questions[1] <- paste0("The *Sum of Squares between* is ***less*** than $",fmt(fls_ssb,3),"$.")
    solutions[1] <- ssb < fls_ssb
    explanations[1] <- paste0("The correct answer is ***",solutions[1],"***. The *Sum of Squares between* is $",fmt(ssb,3),"$, and hence, ",ifelse(solutions[1], "***is***", "is ***not***")," less than $",fmt(fls_ssb,3),"$.")
  } else {
    questions[1] <- paste0("The *Sum of Squares between* is ***greater*** than $",fmt(fls_ssb,3),"$.")
    solutions[1] <- ssb > fls_ssb
    explanations[1] <- paste0("The correct answer is ***",solutions[1],"***. The *Sum of Squares between* is $",fmt(ssb,3),"$, and hence, ",ifelse(solutions[1], "***is***", "is ***not***")," greater than $",fmt(fls_ssb,3),"$.")
  }

# Question 2 (Sum of Squares Within)
if(runif(1) > 0.5) {
    questions[2] <- paste0("The *Sum of Squares within* is ***less*** than $",fmt(fls_ssw,3),"$.")
    solutions[2] <- ssw < fls_ssw
    explanations[2] <- paste0("The correct answer is ***",solutions[2],"***. The *Sum of Squares within* is $",fmt(ssw,3),"$, and hence, ",ifelse(solutions[2], "***is***", "is ***not***")," less than $",fmt(fls_ssw,3),"$.")
  } else {
    questions[2] <- paste0("The *Sum of Squares within* is ***greater*** than $",fmt(fls_ssw,3),"$.")
    solutions[2] <- ssw > fls_ssw
    explanations[2] <- paste0("The correct answer is ***",solutions[2],"***. The *Sum of Squares within* is $",fmt(ssw,3),"$, and hence, ",ifelse(solutions[2], "***is***", "is ***not***")," greater than $",fmt(fls_ssw,3),"$.")
  }

```

```{r ifAllFalseMakeOneTRUE}
# Ensure that at least one answer above is TRUE

i <- 0

if(all(solutions==FALSE)) {
i <- sample.int(2,1)
}

if(i==1) {
  solutions[1] <- TRUE
  if(ssb < fls_ssb) {
  questions[1] <- paste0("The *Sum of Squares between* is ***less*** than $",fmt(fls_ssb,3),"$.")
  explanations[1] <- paste0("The correct answer is ***TRUE***. The *Sum of Squares between* is $",fmt(ssb,3),"$, and hence, ***is*** less than $",fmt(fls_ssb,3),"$.")
  } else {
    questions[1] <- paste0("The *Sum of Squares between* is ***greater*** than $",fmt(fls_ssb,3),"$.")
    explanations[1] <- paste0("The correct answer is ***TRUE***. The *Sum of Squares between* is $",fmt(ssb,3),"$, and hence, ***is*** greater than $",fmt(fls_ssb,3),"$.")
  }
} else if (i==2) {
    solutions[2] <- TRUE
    if(ssw < fls_ssw) {
    questions[2] <- paste0("The *Sum of Squares within* is ***less*** than $",fmt(fls_ssw,3),"$.")
    explanations[2] <- paste0("The correct answer is ***TRUE***. The *Sum of Squares within* is $",fmt(ssw,3),"$, and hence, ***is*** less than $",fmt(fls_ssw,3),"$.")
  } else {
    questions[2] <- paste0("The *Sum of Squares within* is ***greater*** than $",fmt(fls_ssw,3),"$.")
    explanations[2] <- paste0("The correct answer is ***TRUE***. The *Sum of Squares within* is $",fmt(ssw,3),"$, and hence, ***is*** greater than $",fmt(fls_ssw,3),"$.")
  }
}

rm(i)

```

```{r ifAllTrueMakeOneFALSE}
# Ensure that at least one answer above is TRUE

j <- 0

if(all(solutions==TRUE)) {
j <- sample.int(2,1)
}

if(j==1) {
  solutions[1] <- FALSE
  if(ssb > fls_ssb) {
  questions[1] <- paste0("The *Sum of Squares between* is ***less*** than $",fmt(fls_ssb,3),"$.")
  explanations[1] <- paste0("The correct answer is ***FALSE***. The *Sum of Squares between* is $",fmt(ssb,3),"$, and hence, is ***not*** less than $",fmt(fls_ssb,3),"$.")
  } else {
    questions[1] <- paste0("The *Sum of Squares between* is ***greater*** than $",fmt(fls_ssb,3),"$.")
    explanations[1] <- paste0("The correct answer is ***FALSE***. The *Sum of Squares between* is $",fmt(ssb,3),"$, and hence, is ***not*** greater than $",fmt(fls_ssb,3),"$.")
  }
} else if (j==2) {
    solutions[2] <- FALSE
    if(ssw > fls_ssw) {
    questions[2] <- paste0("The *Sum of Squares within* is ***less*** than $",fmt(fls_ssw,3),"$.")
    explanations[2] <- paste0("The correct answer is ***FALSE***. The *Sum of Squares within* is $",fmt(ssw,3),"$, and hence, is ***not*** less than $",fmt(fls_ssw,3),"$.")
  } else {
    questions[2] <- paste0("The *Sum of Squares within* is ***greater*** than $",fmt(fls_ssw,3),"$.")
    explanations[2] <- paste0("The correct answer is ***FALSE***. The *Sum of Squares within* is $",fmt(ssw,3),"$, and hence, is ***not*** greater than $",fmt(fls_ssw,3),"$.")
  }
}

rm(j)

```

```{r QuestionList, echo=F, results = "asis"}
answerlist(questions, markup = "markdown")
```

# Solution

<br/>

The formula for the *Sum of Squares between* is as follows:

$$SS_b=\sum\limits_{k=1}^GN_k(\bar{Y}_k-\bar{Y})^2$$ <br/> Below is a
table that corresponds to the template given in class to solve the *Sum
of Squares between*:

<br/>

|  group $k$  | group mean ($\bar{Y}_k$) | grand mean $\bar{Y}$ | deviation $\bar{Y}_k-\bar{Y}$ | squared deviation $(\bar{Y}_k-\bar{Y})^2$ | group size $N_k$ | weighted squared deviation $N_k(\bar{Y}_k-\bar{Y})^2$ |
|:---------:|:---------:|:---------:|:---------:|:---------:|:---------:|:---------:|
| `r dG$x[1]` |     `r dG$ybark[1]`      |    `r dG$ybar[1]`    |         `r dG$dev[1]`         |              `r dG$devSq[1]`              |   `r dG$n[1]`    |                   `r dG$WdevSq[1]`                    |
| `r dG$x[2]` |     `r dG$ybark[2]`      |    `r dG$ybar[2]`    |         `r dG$dev[2]`         |              `r dG$devSq[2]`              |   `r dG$n[2]`    |                   `r dG$WdevSq[2]`                    |
| `r dG$x[3]` |     `r dG$ybark[3]`      |    `r dG$ybar[3]`    |         `r dG$dev[3]`         |              `r dG$devSq[3]`              |   `r dG$n[3]`    |                   `r dG$WdevSq[3]`                    |
| `r dG$x[4]` |     `r dG$ybark[4]`      |    `r dG$ybar[4]`    |         `r dG$dev[4]`         |              `r dG$devSq[4]`              |   `r dG$n[4]`    |                   `r dG$WdevSq[4]`                    |
|             |                          |                      |                               |                                           |                  |                    $\sum$ `r ssb`                     |

The *Sum of Squares between* is at the lower right in the table above.

------------------------------------------------------------------------

The formula for the *Sum of Squares within* is as follows:

$$SS_w=\sum\limits_{k=1}^G\sum\limits_{i=1}^{N_k}(Y_{ik}-\bar{Y}_k)^2$$
<br/> Below is a table that corresponds to the template given in class
to solve the *Sum of Squares within*:

<br/>

| Group $k$ (degrees Fahrenheit) | Outcome $Y_{i_k}$ | Group mean $\bar{Y}_k$ | Deviation from group mean $Y_{i_k}-\bar{Y}_k$ | Squared deviation $(Y_{i_k}-\bar{Y}_k)^2$ |
|:-------------:|:-------------:|:-------------:|:-------------:|:-------------:|
|           `r d$x[1]`           |    `r d$y[1]`     |     `r d$ybark[1]`     |                 `r d$dev[1]`                  |              `r d$devSq[1]`               |
|           `r d$x[2]`           |    `r d$y[2]`     |     `r d$ybark[2]`     |                 `r d$dev[2]`                  |              `r d$devSq[2]`               |
|           `r d$x[3]`           |    `r d$y[3]`     |     `r d$ybark[3]`     |                 `r d$dev[3]`                  |              `r d$devSq[3]`               |
|           `r d$x[4]`           |    `r d$y[4]`     |     `r d$ybark[4]`     |                 `r d$dev[4]`                  |              `r d$devSq[4]`               |
|           `r d$x[5]`           |    `r d$y[5]`     |     `r d$ybark[5]`     |                 `r d$dev[5]`                  |              `r d$devSq[5]`               |
|           `r d$x[6]`           |    `r d$y[6]`     |     `r d$ybark[6]`     |                 `r d$dev[6]`                  |              `r d$devSq[6]`               |
|           `r d$x[7]`           |    `r d$y[7]`     |     `r d$ybark[7]`     |                 `r d$dev[7]`                  |              `r d$devSq[7]`               |
|           `r d$x[8]`           |    `r d$y[8]`     |     `r d$ybark[8]`     |                 `r d$dev[8]`                  |              `r d$devSq[8]`               |
|           `r d$x[9]`           |    `r d$y[9]`     |     `r d$ybark[9]`     |                 `r d$dev[9]`                  |              `r d$devSq[9]`               |
|          `r d$x[10]`           |    `r d$y[10]`    |    `r d$ybark[10]`     |                 `r d$dev[10]`                 |              `r d$devSq[10]`              |
|          `r d$x[11]`           |    `r d$y[11]`    |    `r d$ybark[11]`     |                 `r d$dev[11]`                 |              `r d$devSq[11]`              |
|          `r d$x[12]`           |    `r d$y[12]`    |    `r d$ybark[12]`     |                 `r d$dev[12]`                 |              `r d$devSq[12]`              |
|                                |                   |                        |                                               |              $\sum$ `r ssw`               |

The *Sum of Squares within* is at the lower right in the table above.

------------------------------------------------------------------------

In case you are interested, below is a boxplot representation of
`Test Score` by `Temperature` group:

```{r, echo=F, warning=F, message=F}
d02 %>% ggplot(aes(Temperature, `Test Score`)) +
  geom_boxplot() +
  geom_point()
```

```{r SolutionList, echo=F, results = "asis"}
answerlist(explanations, markup = "markdown")
```

# Meta-information

extype: mchoice exsolution: `r mchoice2string(solutions)` exname:
SumsOfSquares
