---
title: "Independent samples t-test"
output: html_document
editor_options:
  markdown:
    canonical: false
    wrap: 72
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = F,
	message = FALSE,
	warning = FALSE
)
```

```{r loadPackages, include=FALSE}

# install the packages below if needed
pkgs <- c("truncnorm", "tibble", "dplyr", "forcats", "magrittr", "exams", "kableExtra", "ggplot2")

lapply(pkgs, library, character.only=T)

```

```{r suppressScientificNotation}
options(scipen=999)
```

<!-- Study simulation / Random datasets -->

```{r dataGeneration, echo = FALSE, message=F, warning=F, results = "hide"}
N <- sample(seq(250,300,2),1) # random number between 200 and 250, divisible by 2
propBlack <- c(
  rep("less Black", N/2), 
  rep("more Black", N/2)
  )
Response <- c(
  rtruncnorm(N/2, 1, 6, 3.52, 1.51),
  rtruncnorm(N/2, 1, 6, 3.98, 1.4)
)

d01 <- as_tibble(cbind(propBlack, Response))
d01$ID <- as.factor(rownames(d01))
d01 <- d01 %>% select(c(3,1,2))

d01$Response <- as.numeric(d01$Response)
d01$propBlack <- as.factor(d01$propBlack)

d02 <- d01 %>% 
  rename(`Proportion Black` = propBlack)

s01 <- d02 %>%
  group_by(`Proportion Black`) %>%
  summarize(
    `Mean` = mean(Response),
    `SD` = sd(Response),
    `n` = n()
    )

d03 <- data.frame(
    ID = as.factor(1:(N/2)),
    `less Black` = d02$Response[1:(N/2)],
    `more Black` = d02$Response[(N/2+.1):N]
)

x1 <- round(s01$Mean[1],4)
x2 <- round(s01$Mean[2],4)

mu_diff <- round((x1-x2),4)

s1 <- round(s01$SD[1],4)
s2 <- round(s01$SD[2],4)

s1sq <- round(s1^2,4)
s2sq <- round(s2^2,4)

n1 <- s01$n[1]
n2 <- s01$n[2]

df1 <- n1-1
df2 <- n2-1

df <- (n1+n2)-2

s1sq_df <- round(s1sq*df1,4)
s2sq_df <- round(s2sq*df2,4)

wt_sum_var <- round((s1sq_df+s2sq_df),4)

wt_SE <- round(((wt_sum_var)/(n1+n2-2)),4)

p_adj <- round(((n1+n2)/(n1*n2)),4)

wt_adj <- round((wt_SE*p_adj),4)

pld_SE <- round(sqrt(wt_adj),4)

t <- round((mu_diff/pld_SE),3)

# test <- t.test(Response ~ propBlack, data=d01, alternative = "two.sided", var.equal = TRUE)

```

<!-- Set-up text for student -->

# Question

Hetey and Eberhardt (2014; Study 2) studied the effect of presenting New
York participants two different sets of statistics concerning racial
disparities in the prison population. In one condition (`less-Black`),
they presented participants with accurate national statistics that the
prison population was 40.3% black and 31.8% white. In the `more-Black`
condition, the presented participants with accurate statistics, but
specific to New York City corrections facilities, where the prison
population was approximately 60.3% black and 11.8% white. Then, among
other things, they asked participants a series of questions on a 1-6
Likert scale about how concerned they were about crime (1 = *Not at
all*; 6 = *Extremely*). They found that those in the `more-Black`
condition expressed more concern about crime (*M* = 3.98, *SD* = 1.4)
than those in the `less-Black` condition (*M* = 3.52, *SD* = 1.51),
*t*(162) = -1.998, *p* = .047.

------------------------------------------------------------------------

You replicate this study with `r N` participants. Your results are as
follows:

| `Proportion Black` |     `Mean`      |      `SD`       |  `n`   |
|:------------------:|:---------------:|:---------------:|:------:|
|    `less Black`    | `r round(x1,4)` | `r round(s1,4)` | `r n1` |
|    `more Black`    | `r round(x2,4)` | `r round(s2,4)` | `r n2` |

------------------------------------------------------------------------

**Calculations**

Using the formula for Student's *t* (for independent samples) and using
`less Black` as *x*<sub>1</sub> and `more Black` as *x*<sub>2</sub>,
calculate the degrees of freedom, the obtained *t* (rounded to four
decimal places), and determine whether the difference is statistically
significant. At each stage of your calculation, round to ***FOUR***
decimal places. But your final answer should be rounded to ***THREE***
decimal places.

You will need to make reference too this [lookup table for critical
values of
t](https://patrickabolger.github.io/PSYC-301-Exam-Formulas/#t-table)

------------------------------------------------------------------------

**TRUE/FALSE Questions**

Now that you have calculated these statistics, answer the questions
below. Check the box to indicate **TRUE**, and leave it blank to
indicate **FALSE**.

------------------------------------------------------------------------

<br/>

<!-- TRUE/FALSE (mchoice) question randomization -->

```{r QuestionVectorSize, echo=F, results = "hide"}
questions <-    character(4)
solutions <-    logical(4)
explanations <- character(4)
```

```{r QuestionGeneration, echo=F, results = "hide"}

# t-critical
tcrit <- abs(round(qt(.025, df, lower.tail=TRUE),3))

# p-value
p <- 2*pt(t, df)

fls_df <- sample(c(N-4,N-3,N-1,N,N+1,N+2,N+3), 1)
fls_t <- round(sample(c(t - .05*t, t + .05*t),1), 3)
fls_tcrit <- round(sample(c(tcrit - .01*tcrit, tcrit + .01*tcrit),1), 3)

# Question 1 (df)
if(runif(1) > 0.5) {
  questions[1] <- paste0("The **degrees of freedom** are **less** than $",fmt(fls_df,0),"$.")
  solutions[1] <- df < fls_df
  explanations[1] <- paste0("The correct answer is ",solutions[1],". The degrees of freedom are $",fmt(df,0),"$, and hence, ",ifelse(solutions[1], "**are**", "are **not**")," less than $",fmt(fls_df,0),"$.")
} else {
  questions[1] <- paste0("The **degrees of freedom** are **greater** than $",fmt(fls_df,0),"$.")
  solutions[1] <- df > fls_df
  explanations[1] <- paste0("The correct answer is ",solutions[1],". The degrees of freedom are $",fmt(df,0),"$, and hence, ",ifelse(solutions[1], "**are**", "are **not**")," greater than $",fmt(fls_df,0),"$.")
}

# Question 2 (obtained t)
if(runif(1) > 0.5) {
    questions[2] <- paste0("The obtained value for *t* is **less** than $",fmt(fls_t,3),"$.")
    solutions[2] <- t < fls_t
    explanations[2] <- paste0("The correct answer is ",solutions[2],". The obtained value for *t* is $",fmt(t,3),"$, and hence, ",ifelse(solutions[2], "**is**", "is **not**")," less than $",fmt(fls_t,3),"$.")
  } else {
    questions[2] <- paste0("The obtained value for *t* is **greater** than $",fmt(fls_t,3),"$.")
    solutions[2] <- t > fls_t
    explanations[2] <- paste0("The correct answer is ",solutions[2],". The obtained value for *t* is $",fmt(t,3),"$, and hence, ",ifelse(solutions[2], "**is**", "is **not**")," greater than $",fmt(fls_t,3),"$.")
  }

# Question 3 (critical value for t)
if(runif(1) > 0.5) {
    questions[3] <- paste0("The critical value (two-tailed, alpha = .05) for *t* is **less** than $",fmt(fls_tcrit,3),"$.")
    solutions[3] <- tcrit < fls_tcrit
    explanations[3] <- paste0("The correct answer is ",solutions[3],". The critical value (two-tailed, alpha = .05) for *t* with $",fmt(df,0),"$ degrees of freedom is $",fmt(tcrit,3),"$, and hence, ",ifelse(solutions[3], "**is**", "is **not**")," less than $",fmt(fls_tcrit,3),"$.")
  } else {
    questions[3] <- paste0("The critical value (two-tailed, alpha = .05) for *t* is **greater** than $",fmt(fls_tcrit,3),"$.")
    solutions[3] <- tcrit > fls_tcrit
    explanations[3] <- paste0("The correct answer is ",solutions[3],". The critical value (two-tailed, alpha = .05) for *t* with $",fmt(df,0),"$ degrees of freedom is $",fmt(tcrit,3),"$, and hence, ",ifelse(solutions[3], "**is**", "is **not**")," greater than $",fmt(fls_tcrit,3),"$.")
  }

# Question 4 (significance)
questions[4] <- paste0("The results of the *t*-test (two-tailed) are statistically significant.")
solutions[4] <- abs(t) >= tcrit
explanations[4] <- paste0("The correct answer is ",solutions[4],". The absolute value of the obtained value for *t* is $",fmt(abs(t),3),"$, and hence, ",ifelse(solutions[4], "**is**", "is **not**")," greater than or equal to the critical value (two-tailed) of $",fmt(tcrit,3),"$.")

```

```{r iifAllFalseMakeOneTRUE, echo=F, results = "hide"}
i <- 0

if(all(solutions==FALSE)) {
  i <- sample.int(4,1)
  }

if(i==1) {
  solutions[1] <- TRUE
  if(df < fls_df) {
  questions[1] <- paste0("The **degrees of freedom** are **less** than $",fmt(fls_df,0),"$.")
  explanations[1] <- paste0("The correct answer is *TRUE*. The degrees of freedom are $",fmt(df,0),"$, and hence, are less than $",fmt(fls_df,0),"$.")
  } else {
    questions[1] <- paste0("The **degrees of freedom** are **greater** than $",fmt(fls_df,0),"$.")
    explanations[1] <- paste0("The correct answer is *TRUE*. The degrees of freedom are $",fmt(df,0),"$, and hence, are greater than $",fmt(fls_df,0),"$.")
  }
} else if (i==2) {
    solutions[2] <- TRUE
    if(t < fls_t) {
    questions[2] <- paste0("The obtained value for *t* is **less** than $",fmt(fls_t,3),"$.")
    explanations[2] <- paste0("The correct answer is *TRUE*. The obtained value for *t* is $",fmt(t,3),"$, and hence, is less than $",fmt(fls_t,3),"$.")
  } else {
    questions[2] <- paste0("The obtained value for *t* is **greater** than $",fmt(fls_t,3),"$.")
    explanations[2] <- paste0("The correct answer is *TRUE*. The obtained value for *t* is $",fmt(t,3),"$, and hence, is greater than $",fmt(fls_t,3),"$.")
  }
} else if (i==3) {
    solutions[3] <- TRUE
    if(tcrit < fls_tcrit) {
    questions[3] <- paste0("The critical value (two-tailed, alpha = .05) for *t* is **less** than $",fmt(fls_tcrit,3),"$.")
    explanations[3] <- paste0("The correct answer is *TRUE*. The critical value (two-tailed, alpha = .05) for *t* with $",fmt(df,0),"$ degrees of freedom is $",fmt(tcrit,3),"$, and hence, is less than $",fmt(fls_tcrit,3),"$.")
  } else {
    questions[3] <- paste0("The critical value (two-tailed, alpha = .05) for *t* is **greater** than $",fmt(fls_tcrit,3),"$.")
    explanations[3] <- paste0("The correct answer is *TRUE*. The critical value (two-tailed, alpha = .05) for *t* with $",fmt(df,0),"$ degrees of freedom is $",fmt(tcrit,3),"$, and hence, is greater than $",fmt(fls_tcrit,3),"$.")
  }
} else if (i==4) {
  solutions[4] <- TRUE
  if(abs(t) >= tcrit) {
      questions[4] <- paste0("The results of the *t*-test (two-tailed) are statistically significant.")
      explanations[4] <- paste0("The correct answer is TRUE. The absolute value of the obtained value for *t* is $",fmt(abs(t),3),"$, and hence, is greater than or equal to the critical value (two-tailed) of $",fmt(tcrit,3),"$.")
  } else {
      questions[4] <- paste0("The results of the *t*-test (two-tailed) are **not** statistically significant.")
      explanations[4] <- paste0("The correct answer is TRUE. The absolute value of the obtained value for *t* is $",fmt(abs(t),3),"$, and hence, is less than the critical value (two-tailed) of $",fmt(tcrit,3),"$.")
  }
}

rm(i)
```

```{r ifAllTrueMakeOneFalse, echo=F, results = "hide"}
j <- 0

if(all(solutions==TRUE)) {
  j <- sample.int(4,1)
  }

if(j==1) {
  solutions[1] <- FALSE
  if(df > fls_df) {
  questions[1] <- paste0("The **degrees of freedom** are **less** than $",fmt(fls_df,0),"$.")
  explanations[1] <- paste0("The correct answer is *FALSE*. The degrees of freedom are $",fmt(df,0),"$, and hence, are greater than $",fmt(fls_df,0),"$.")
  } else {
    questions[1] <- paste0("The **degrees of freedom** are **greater** than $",fmt(fls_df,0),"$.")
    explanations[1] <- paste0("The correct answer is *FALSE*. The degrees of freedom are $",fmt(df,0),"$, and hence, are less than $",fmt(fls_df,0),"$.")
  }
} else if (j==2) {
    solutions[2] <- FALSE
    if(t > fls_t) {
    questions[2] <- paste0("The obtained value for *t* is **less** than $",fmt(fls_t,3),"$.")
    explanations[2] <- paste0("The correct answer is *FALSE*. The obtained value for *t* is $",fmt(t,3),"$, and hence, is greater than $",fmt(fls_t,3),"$.")
  } else {
    questions[2] <- paste0("The obtained value for *t* is **greater** than $",fmt(fls_t,3),"$.")
    explanations[2] <- paste0("The correct answer is *FALSE*. The obtained value for *t* is $",fmt(t,3),"$, and hence, is less than $",fmt(fls_t,3),"$.")
  }
} else if (j==3) {
    solutions[3] <- FALSE
    if(tcrit > fls_tcrit) {
    questions[3] <- paste0("The critical value (two-tailed, alpha = .05) for *t* is **less** than $",fmt(fls_tcrit,3),"$.")
    explanations[3] <- paste0("The correct answer is *FALSE*. The critical value (two-tailed, alpha = .05) for *t* with $",fmt(df,0),"$ degrees of freedom is $",fmt(tcrit,3),"$, and hence, is greater than $",fmt(fls_tcrit,3),"$.")
  } else {
    questions[3] <- paste0("The critical value (two-tailed, alpha = .05) for *t* is **greater** than $",fmt(fls_tcrit,3),"$.")
    explanations[3] <- paste0("The correct answer is *FALSE*. The critical value (two-tailed, alpha = .05) for *t* with $",fmt(df,0),"$ degrees of freedom is $",fmt(tcrit,3),"$, and hence, is less than $",fmt(fls_tcrit,3),"$.")
  }
} else if (j==4) {
  solutions[4] <- FALSE
  if(abs(t) < tcrit) {
      questions[4] <- paste0("The results of the *t*-test (two-tailed) are statistically significant.")
      explanations[4] <- paste0("The correct answer is FALSE The absolute value of the obtained value for *t* is $",fmt(abs(t),3),"$, and hence, is less than the critical value (two-tailed) of $",fmt(tcrit,3),"$.")
  } else {
      questions[4] <- paste0("The results of the *t*-test (two-tailed) are **not** statistically significant.")
      explanations[4] <- paste0("The correct answer is FALSE The absolute value of the obtained value for *t* is $",fmt(abs(t),3),"$, and hence, is greater than or equal to the critical value (two-tailed) of $",fmt(tcrit,3),"$.")
  }
}


rm(j)
```

```{r QuestionList, echo=F, results = "asis"}
answerlist(questions, markup = "markdown")
```

<!-- Solutions -->

# Solution

The results of this *t*-test are as follows:\
<br/>

$$t(`r df`)= `r round(t,3)`, p = `r p`$$

---
The degrees of freedom for a student's (independent-samples) *t*-test are as follows:
Thus, for the current data set, they are calculated as follows:
editor_options: 
  markdown: 
    wrap: 72
---

<br/>

The obtained value for Student's *t* (independent-samples) is calculated
with the following formula:

$$t=\frac{\overline{x}_1-\overline{x}_2}{\sqrt{\bigg[\frac{(n_1-1)s_1^2+(n_2-1)s_2^2}{n_1+n_2-2}\bigg]\bigg[\frac{n_1+n_2}{n_1n_2}\bigg]}}$$
<br/>

So the obtained value for Student's *t* is as follows: <br/>

$$=\frac{`r x1`-`r x2`}{\sqrt{\bigg[\frac{(`r n1`-1)`r s1`^2+(`r n2`-1)`r s2`^2}{`r n1`+`r n2`-2}\bigg]\bigg[\frac{`r n1`+`r n2`}{`r n1`\times`r n2`}\bigg]}}$$

<br/>

$$=\frac{`r mu_diff`}{\sqrt{\bigg[\frac{(`r df1`)`r s1sq`+(`r df1`)`r s2sq`}{`r n1+n2`-2}\bigg]\bigg[\frac{`r n1+n2`}{`r n1*n2`}\bigg]}}$$

<br/>

$$=\frac{`r mu_diff`}{\sqrt{\bigg[\frac{`r s1sq_df`+`r s2sq_df`}{`r df`}\bigg]\bigg[`r p_adj`\bigg]}}$$
<br/>
$$=\frac{`r mu_diff`}{\sqrt{\bigg[\frac{`r wt_sum_var`}{`r df`}\bigg]\bigg[`r p_adj`\bigg]}}$$

<br/>
$$=\frac{`r mu_diff`}{\sqrt{\bigg[`r wt_SE`\bigg]\bigg[`r p_adj`\bigg]}}$$

<br/> $$=\frac{`r mu_diff`}{\sqrt{`r wt_adj`}}$$

<br/> $$=\frac{`r mu_diff`}{`r pld_SE`}$$

<br/> $$=`r t`$$

<br/>

To get the critical value (two-tailed) for *t* with $\alpha=.05$, you go
to <https://www.omnicalculator.com/statistics/critical-value> and make
the following choices:

-   What distribution?

    -   *t*\

-   What type of of test?

    -   Two-tailed\

-   Degrees of freedom (d)

    -   `r round(df,0)`\

-   Significance level

    -   0.05

-   Then look below those settings for the critical value

    -   You should see $t_{critical}(`r df`)=`r round(tcrit,3)`$

------------------------------------------------------------------------

*Reference*

Hetey, R. C., & Eberhardt, J. L. (2014). Racial disparities in
incarceration increase acceptance of punitive policies. *Psychological
Science*, *25*(10), 1949-1954.

------------------------------------------------------------------------

<!-- Explanations -->

```{r SolutionList, echo=F, results = "asis"}
answerlist(explanations, markup = "markdown")
```

<!-- Metadata -->

# Meta-information

extype: mchoice exsolution: `r mchoice2string(solutions)` exname: t-test
